---
title: "20180310_GB_A6 and Adducts"
author: "Garrett Boudinot"
date: "`r format(Sys.Date(), '%d %b %Y')`"
output:
  html_document: 
    df_print: paged
    toc: yes
    toc_float: true
    toc_depth: 3
    code_folding: show
editor_options:
  chunk_output_type: inline
---

# Normal Start Functions

```{r "cleanup", echo = FALSE, eval=FALSE}
# not run in knitted doc
rm(list = ls())
.rs.restartR()
```

```{r setup, echo = FALSE, message=FALSE, warning=FALSE}
# load libraries
library(tidyverse) 
library(plotly)
library(isoreader)
library(dplyr)
library(ggplot2)
library(knitr)
library(readxl)
library(tidyr)
library(isoprocessorCUB)
source("functions.R")
# set global chunk options for figure output
figure_prefix <- "PREFIX-"
knitr::opts_chunk$set(
  dev=c("png", "pdf"), dev.args=list(pdf = list(encoding="WinAnsi", useDingbats=FALSE)),
  fig.keep="all", fig.path=file.path("figures", figure_prefix))
```

# Load isofiles
```{r, message=FALSE}
isofiles <-
  file.path(
    "Results",
    c(
      "180305_alkanes_C_GB",
      "180306_alkanes_C_GB",
      "180307_alkanes_C_GB",
      "180309_alkanes_C_GB"
    )
  ) %>%
  iso_read_continuous_flow()
```

# Data

## File Info
```{r}
# all file info
isofiles %>% iso_get_file_info()
```

## Vendor data table
```{r}
isofiles %>% iso_get_vendor_data_table()
```

## Chromatograms
```{r, fig.width=10, fig.height=8, message=FALSE}
isofiles %>% 
  # fetch a few of the files of interst
  iso_filter_files(parse_number(Analysis) %in% c(616, 632)) %>% 
  # plot just mass 28
  iso_plot_continuous_flow_data(data = c(46)) %>% 
  # make interactive
  ggplotly(dynamicTicks = TRUE)
```

# Analysis

## Select relevant data
```{r}
data_table <-
  isofiles %>%
  # filter the files you want to use 
  # --> exclude CO2 zeros
  iso_filter_files(!str_detect(`Identifier 1`, "CO2 zero")) %>% 
  # --> select only good analyses
  iso_filter_files(parse_number(Analysis) > 572) %>% 
  # get the vendor data table and file info
  iso_get_vendor_data_table(
    select = c(
      # peak info
      Nr., Start, Rt, End,
      # amplitudes and intensities
      `Ampl 44`, `Ampl 46`, `Intensity All`, `Intensity 44`,
      # ratios and deltas
      `R 46CO2/44CO2`, `d 13C/12C`
    ),
    include_file_info = c(
      file_datetime, Analysis, `Identifier 1`, `Identifier 2`, `GC Method`, `Seed Oxidation`
    )
  ) %>% 
  # rename some columns to be easier to work with
  rename(
    Ampl44=`Ampl 44`, Ampl46=`Ampl 46`, # amplitudes
    #Area28=`rIntensity 28`, Area29=`rIntensity 29`, # areas
    Intensity=`Intensity All`, #peak intensities
    R46C = `R 46CO2/44CO2`, d13C = `d 13C/12C`, # ratio and delta values
    rt = `Rt`,
    rt_start = `Start`,
    rt_end = `End`
    #File = `file_id`
  )
data_table
```

## Map peaks
```{r, echo=FALSE, eval=FALSE}
# file names
data_table$file_id %>% unique() %>%  paste(collapse = "\n") %>% message()
```

```{r}
### CHANGE MAPPING FILE NAME ###
metadata_samples <- read_excel(file.path("metadata", "20180310_GB_metadata.xlsx"), sheet = "files")

metadata_peak_maps <- read_excel(file.path("metadata","20180310_GB_metadata.xlsx"), sheet = "maps")

metadata_samples
```

```{r}
data_table_with_peaks1 <- data_table %>%
    iso_add_metadata(metadata_samples, match_by = c(file_id)) %>%
    iso_map_peaks(metadata_peak_maps) 

# missing and unidentified peaks
data_table_with_peaks1 %>% filter(!is_identified | is_missing) %>% filter(process == "yes")

# confirmed peaks
data_table_with_peaks1 <- data_table_with_peaks1 %>%
    filter(is_identified, !is_missing, !is_ambiguous)

# missing and unidentified peaks
data_table_with_peaks <- data_table_with_peaks1 %>% filter(process == "yes")

data_table_with_peaks
```



## Check stability of reference peaks
The second one is the one defined to be 0 permil (so will always be), the rest is relative to that peak.
```{r, message=FALSE, fig.width = 10, fig.height = 8}
p <- data_table_with_peaks %>% 
  filter(is_ref_peak == "yes") %>% 
  ggplot() + 
  aes(Nr., d13C, color = file_id) +
  geom_line() + 
  theme_bw()
ggplotly(p)
```

## Add standard values
```{r}
standards <- read_excel(file.path("metadata", "gc_irms_indiana_A6.xlsx"))
data_w_stds <- data_table_with_peaks %>% 
  filter(type == "standard", is_ref_peak == "no") %>% 
  left_join(standards, by = "compound") %>% 
  mutate(is_std = !is.na(true.d13C) | !is.na(true.d2H))
```


# Process data
Focus on the analytes and calculate a few summary parameters we want to use later.
```{r}
data_w_analyte_peaks <- 
  data_table_with_peaks %>% 
  # this is important so that the reference peaks are not caught up in the next set of calculations
  filter(is_ref_peak == "no") %>% 
  # for each analysis calculate averages across analysis
  group_by(file_id) %>% 
  mutate(
    ampl_sample_mean.mV = mean(`Ampl44`), ampl_sample_sd.mV = sd(`Ampl44`),
    area_sample_mean.Vs = mean(`Intensity 44`), area_sample_sd.Vs = sd(`Intensity 44`)
  )
```

## Standards
```{r}
standards <- read_excel(file.path("metadata", "gc_irms_indiana_A6.xlsx"))
kable(standards)

data_w_stds1 <-
  data_w_analyte_peaks %>% 
  iso_add_standards(standards, "compound") 

data_w_stds <- data_w_stds1 %>% filter(type == "standard") 
```

### Visualize standards
```{r, fig.width = 9, fig.height=6}
v <- data_w_stds %>% 
  ggplot() +
  aes(x = true.d13C, y = d13C, color = file_id) + 
  geom_smooth(method = "lm", se = FALSE, alpha = 0.5) +
  geom_point() +
  theme_bw() +
  theme(legend.position = "none") 
ggplotly(v)
```

## Calibration
```{r}
data_w_calibs <- data_w_stds %>% 
  # prepare for calibration by defining the grouping column(s) and setting default parameters
  iso_prepare_for_calibration(group_by = c(file_id)) %>% 
  iso_set_default_process_parameters(delta_residual = resid_d13C) %>% 
  # run calibration
  iso_generate_calibration(model = lm(`d13C` ~ true.d13C)) #%>% 
  # pull out some columns we want generally available
 # iso_unnest_calibration_parameters(select == ampl_sample_mean.mV) 

data_w_calibs %>% 
  iso_unnest_calibration_summary(keep_other_list_data = FALSE) %>% 
  kable(digits =3)

data_w_calibs %>% 
  iso_unnest_calibration_coefs(select = c(-statistic), keep_other_list_data = FALSE) %>% 
  # arrange by term and Analysis to get a quick idea of what the numbers across analyses
  arrange(term, file_id) %>% 
  kable(digits = 2)
```

### Parameters
```{r "calibration_parameters", fig.width = 8, fig.height = 8}
data_params <- data_w_calibs %>% 
  # pull out remaining columns we want available (some might already be pulled out but that's okay)
  #Note: Garrett changed "Preparation" to "Seed Oxidation" - we may want a different shape variable
  iso_unnest_calibration_range(select = c(file_datetime, ampl_sample_mean.mV, `Seed Oxidation`, is_standard)) 
  
# visualize the delta calibration fits
data_params %>% 
  #NEED TO FILTER OUT SAMPLES iso_filter_files(`Identifier 1` == "A5") 
  iso_visualize_delta_calib_fits(x = file_id, color = `Identifier 2`, size = ampl_sample_mean.mV,
                                 include_from_summary = c(adj.r.squared, deviance)) + labs(title = "parameters vs. analysis")

data_params %>% 
  iso_visualize_delta_calib_fits(x = ampl_sample_mean.mV, color = `Identifier 2`, size = ampl_sample_mean.mV,
                                 include_from_summary = c(adj.r.squared, deviance)) + labs(title = "parameters vs. amplitude") 

data_params %>% 
  iso_visualize_delta_calib_fits(x = file_datetime, color = `Identifier 2`, size = ampl_sample_mean.mV,
                                 include_from_summary = c(adj.r.squared, deviance)) + labs(title = "parameters vs. time")
```

```{r "interactive_calibration_parameters", fig.width = 8, fig.height = 8}
# turn the last plot into an interactive one
ggplotly(ggplot2::last_plot() + theme(legend.position = "none"))
```

#### Compounds
```{r "standards_by_compound", fig.width = 8, fig.height = 8}
data_w_calibs %>% 
  # pull out relevant data
  iso_unnest_calib_data(select = everything()) %>% 
  filter(is_standard) %>% 
  # calculate deviation from means
  group_by(file_id) %>% 
  mutate(
    `Var: residual d13C [permil]` = resid_d13C,
    `Var: area diff from mean [%]` = (`Intensity 44`/mean(`Intensity 44`) - 1) * 100,
    `Var: amplitude diff from mean [%]` = (`Ampl44`/mean(`Ampl44`) - 1) * 100
  ) %>%
  # visualize
  iso_visualize_data(x = compound, y = starts_with("Var"), group = file_id, color = `Identifier 2`)
```

```{r "interactive", fig.width = 8, fig.height = 8}
ggplotly(ggplot2::last_plot())
```

# Apply correction to Adducts

```{r}
std_corrections <- 
  data_w_stds %>% 
  group_by(file_id) %>% 
  do({
    m <- lm(d13C ~ true.d13C, data = .)
    data_frame(
      offset = coefficients(m)[1],
      slope = coefficients(m)[2],
      delta_ref_CO = -offset/slope,
      max_residual = max(summary(m)$residuals),
      rss = sum(summary(m)$residuals^2),
      r2 = summary(m)$r.squared
    )
  }) 
std_corrections %>% knitr::kable(d = 3)
```

```{r}
message(sprintf("Infered reference tank isotopic composition: %.2f +/- %.2f",
                std_corrections$delta_ref_CO %>% mean(),
                std_corrections$delta_ref_CO %>% sd()))

# simplifed offset=ref gas correction (TODO: consider signal strengh (linearity correct), consider drift and evaluate standards between samples rather than average)
offset_avg <- std_corrections$offset %>% mean()
slope_avg <- std_corrections$slope %>% mean()


final_data <- data_w_stds1 %>% 
  filter(type == "sample", is_ref_peak == "no") %>% 
  mutate(d13C.corrected = (d13C - offset_avg)/(slope_avg))
```

#### All

```{r}
final_data %>% 
  ggplot() +
  aes(compound, y = d13C.corrected, color = file_id) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90))
```

```{r}
ggplotly(ggplot2::last_plot() + theme(legend.position = "none"))
```

#### Against depth

```{r "overview_against_depth", fig.width = 10, fig.height = 8}
subset(final_data, compound %in% c("C31", "C27")) %>% 
  ggplot() +
  aes(x = Depth, y = d13C.corrected, color = file_id, size = Ampl44) +
  geom_point() +
  scale_x_reverse() + 
  facet_wrap(~compound, scales = "free") + 
  coord_flip()
```

```{r}
ggplotly(ggplot2::last_plot() + theme(legend.position = "none"))
```

```{r}
#saveRDS(Cache, file = "data_w_stds1")
saveRDS(final_data, file = file.path("cache", "dataset_march2018_corrected.rds"))

```

```{r}
final_data %>%
      filter(compound == "C31") %>% print(compound, everything())
```

