---
title: "20180720_GB_SH1_d13Call_presentation"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

```{r "cleanup", echo = FALSE, eval=FALSE}
# not run in knitted doc
rm(list = ls())
.rs.restartR()
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}

library(isoprocessor)
library(readxl)
library(tidyverse)
library(plotly)
library(knitr)
library(isoreader)
library(gridExtra)
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```


# Load data
```{r, echo = FALSE}
# load dataset
iso_files <- iso_read_continuous_flow("Garrett_Adducts.cf.rda")
```

## Vendor data table

```{r, echo = FALSE}
vdt <- iso_files %>% 
  # get relevant vendor data table columns and file info
  iso_get_vendor_data_table(
    select = c(starts_with("Amp"), everything()),
    include_file_info = c(ox = `Seed Oxidation`, everything())
  )
```

## Map peaks
```{r, echo=FALSE, eval=FALSE}
# file names
vdt$file_id %>% unique() %>%  paste(collapse = "\n") %>% message()
```

## Chromatograms
```{r, echo = FALSE, fig.width=10, fig.height=8, message=FALSE}
iso_files %>% 
  # fetch a few of the files of interst
  iso_filter_files(parse_number(Analysis) %in% c(603, 595)) %>% 
  # plot just mass 28
  iso_plot_continuous_flow_data(data = c(46)) %>% 
  # make interactive
  ggplotly(dynamicTicks = TRUE)
```

## Retrieve metadata

```{r, echo = FALSE}
metadata_samples <- read_excel(file.path("metadata", "20180416_GB_metadata.xlsx"), sheet = "files")
metadata_peak_maps <- read_excel(file.path("metadata","20180416_GB_metadata.xlsx"), sheet = "maps")
metadata_samples
```

## Map peaks

```{r, results='asis'}
# add metadata
data_w_metadata <- vdt %>% 
  iso_add_metadata(metadata_samples, match_by = c(`Identifier 1`)) 

# missing metadata
data_w_metadata %>% 
  iso_get_missing_metadata(select = c(Analysis, `Identifier 1`, map_id))

# map peaks
data_w_peaks_all <- data_w_metadata %>% 
  # only continue with records that have metadata and are flagged for processing
  iso_remove_missing_metadata() %>% 
  filter(process == "yes") %>% 
  # map peaks
  iso_map_peaks(metadata_peak_maps, file_id = file_id, rt = Rt, rt_start = Start, rt_end = End)

# problematic peaks
prob <- data_w_peaks_all %>% 
  iso_get_problematic_peaks(select = c(file_id, Analysis, compound, Start, Rt, End))

# focus on non problematic peaks only?
data_w_peaks <- data_w_peaks_all %>% iso_remove_problematic_peaks()
data_w_peaks
```

## Reference peaks

```{r "ref_variation", fig.width=8, fig.height=6}
data_w_peaks %>% 
  iso_plot_ref_peaks(
    is_ref_condition = is_ref_peak == "yes", 
    ratio = c(`R 13C/12C`, `R 18O/16O`),
    group_id = `Analysis`,
    within_group = TRUE
  ) 
```


# Data processing

## Focus on analytes 

Focus on the analytes and calculate a few summary parameters we want to use later.

```{r, echo = FALSE}
data_w_analyte_peaks <- 
  data_w_peaks %>% 
  # this is important so that the reference peaks are not caught up in the next set of calculations
  filter(is_ref_peak == "no") %>% 
  # for each analysis calculate averages across analysis
  group_by(file_id) %>% 
  mutate(
    ampl_sample_mean.mV = mean(`Ampl 44`), ampl_sample_sd.mV = sd(`Ampl 44`),
    area_sample_mean.Vs = mean(`Intensity 44`), area_sample_sd.Vs = sd(`Intensity 44`)
  )
```

## Evaluate Standards

### Known isotope values

Add known isotope values for standards.

```{r, echo = FALSE}
standards <- 
  read_excel(file.path("metadata", "GC-IRMS standards.xlsx")) %>% 
  mutate(type = "standard") %>% 
  mutate(compound = str_replace(compound, "-", ""))
standards

data_w_stds <-
  data_w_analyte_peaks %>% 
  iso_add_standards(standards, match_by = c(type, compound))  
data_w_stds
```

### Single analysis calibration (for QA)

Determine calibrations fits for individual standard analyses.

```{r, echo = FALSE}
stds_w_calibs <- data_w_stds %>%
  # focus on standard analyses
  filter(type == "standard") %>%  
  # prepare for calibration by defining the grouping column(s) 
  iso_prepare_for_calibration(group_by = c(Analysis)) %>% 
  # pull out additional ID columns from the nested data
  iso_unnest_data(
    select = c(`Analysis`, `Injection Volume` = `Identifier 2`)) %>% 
  # run calibration
  iso_generate_calibration(
    model = lm(`d 13C/12C` ~ true_d13C), 
    calibration = "d13C",
    is_standard = is_standard,
    # only consider the standards that actually have the 8 peaks
    min_n_datapoints = 8
  )

# check for problematic calibrations and make note of them in the main dataset
problematic_stds <- stds_w_calibs %>% iso_get_problematic_calibrations("d13C")
problematic_stds
data_w_stds <- data_w_stds %>% 
  mutate(type = ifelse(Analysis %in% problematic_stds$Analysis,
                       "excluded standard", type))

# move forward only with good calibrations
stds_w_calibs <- stds_w_calibs %>% iso_remove_problematic_calibrations("d13C")
```

### Calibration coefficients

Look at calibration parameters (coefficients and summary) as well ad data ranges in overview table.

```{r, echo = FALSE}
# look at coefficients and summary
stds_w_calibs %>% 
  # pull out a little more file information
  iso_unnest_data(select = c(`Seed Oxidation`)) %>% 
  # unnest calibration parameters
  iso_unnest_calibration_parameters(
    calibration = "d13C", 
    select_from_coefs = c(term, estimate, SE = std.error, signif),
    select_from_summary = c(fit_R2 = adj.r.squared, fit_RMSD = deviance, residual_df = df.residual)) %>% 
  arrange(term)

# look at ranges
stds_w_calibs %>% 
  iso_unnest_calibration_range(
    calibration = "d13C",
    select = c(var, min, max)
  ) %>% 
  arrange(var)
```

Visualize the calibration parameters

```{r "calibration_parameters", fig.width = 8, fig.height = 8}
# unnest some of the data for plotting
data_for_plots <- 
  stds_w_calibs %>%
  iso_unnest_data(
    c(Oxidation = `Seed Oxidation`, 
      datetime = file_datetime,
      `Mean Amplitude` = ampl_sample_mean.mV)
  ) 
data_for_plots


# parameters vs time
data_for_plots %>%
  iso_plot_calibration_parameters(
    calibration = "d13C", 
    x = datetime,
    color = `Injection Volume`,
    shape = Oxidation
  )

# parameter vs analysis
data_for_plots %>% 
  iso_plot_calibration_parameters(
    calibration = "d13C", 
    x = paste(Analysis),
    color = `Injection Volume`,
    shape = Oxidation
  )

# parameters vs amplitude
data_for_plots %>% 
  iso_plot_calibration_parameters(
    calibration = "d13C", 
    x = `Mean Amplitude`,
    color = `Injection Volume`
  ) 
```


```{r "interactive_calibration_parameters", fig.width = 8, fig.height = 8}
# turn the last plot into an interactive one
ggplotly(ggplot2::last_plot() + theme(legend.position = "none"))
```

### Inspect Residuals

Look at residuals to see goodness of fit for the single analysis calibrations.

```{r "standards_by_compound", fig.width = 8, fig.height = 8}
stds_w_calibs %>% 
  # pull out all data
  iso_unnest_data(select = everything()) %>% 
  filter(is_standard) %>% 
  # calculate deviation from means
  group_by(Analysis) %>% 
  mutate(
    `Var: residual d13C [permil]` = d13C_resid,
    `Var: area diff from mean [%]` = (`Intensity 44`/mean(`Intensity 44`) - 1) * 100,
    `Var: amplitude diff from mean [%]` = (`Ampl 44`/mean(`Ampl 44`) - 1) * 100
  ) %>%
  ungroup() %>% 
  # visualize
  iso_plot_data(
    x = compound,
    y = starts_with("Var"),
    group = Analysis,
    color = `Injection Volume`
  ) +
  # plot modifications
  labs(x = "")
```

```{r "interactive", fig.width = 8, fig.height = 8}
ggplotly()
```

## Gobal calibration

### Test different calibration models

Run global calibrations across all standards.

```{r, echo = FALSE}
#original
data_w_calibs <- data_w_stds %>% 
  # no grouping this time, everything is part of the global calibration
  iso_prepare_for_calibration() %>% 
  # run calibrations
  iso_generate_calibration(
    model = c(
      # simple model
      delta_only = lm(`d 13C/12C` ~ true_d13C),
      # multivariate with delta and amplitude (could also check Intensity 44)
      delta_and_ampl = lm(`d 13C/12C` ~ true_d13C + `Ampl 44`),
      # + the delta and amplitude cross term
      delta_cross_ampl = lm(`d 13C/12C` ~ true_d13C * `Ampl 44`),
      # multivariate with delta and the datetime (i.e. checking for temporal drift)
      delta_and_time = lm(`d 13C/12C` ~ true_d13C + file_datetime)
    ), 
    calibration = "d13C",
    # important to identify which analyses and which peaks within those should
    # be part of the calibration
    is_standard = type == "standard" & is_standard
  ) %>% 
  iso_remove_problematic_calibrations("d13C")
```


Visualize the calibration parameters for the different models. 

```{r "global_calibration_parameters", fig.width = 7, fig.height = 10}
# coefficient summary table
data_w_calibs %>% iso_unnest_calibration_parameters("d13C")

# plot to look at coefficients and summary
data_w_calibs %>% 
  iso_plot_calibration_parameters(
    calibration = "d13C",
    x = d13C_calib,
    color = d13C_calib,
    # include the significance level to gauge which model parameters matter
    shape = signif
  )
```


Garrett thinks delta and aplitude and delta cross amplitude look better

```{r "global_calibration_residuals", fig.width=8, fig.height=6}
data_w_calibs %>% 
  iso_unnest_data(select = everything()) %>% 
  rename(`Injection Volume` = `Identifier 2`) %>% 
  filter(type == "standard" & is_standard) %>% 
  # plot each standard analysis for each of the models to see the differences
  iso_plot_data(
    x = compound,
    y = d13C_resid,
    group = paste(Analysis, d13C_calib),
    color = d13C_calib
  ) 
```


### Apply calibration

Apply the `delta_and_ampl` calibration. 

```{r, echo = FALSE}
data_calibrated <- data_w_calibs %>% 
  # apply the delta_and_ampl calibration (could apply multiple and compare)
  filter(d13C_calib == "delta_and_ampl") %>% 
  iso_apply_calibration(
    predict = true_d13C,
    calibration = "d13C",
    calculate_error = TRUE
  ) %>% 
  # unnest the data
  iso_unnest_data()
data_calibrated
```

### Inspect results

The comparison of the predicted value in the standards along with the standard error in the prediction (the inherent error of the global calibration itself, which should be treated as the analytical error) can give a better sense of how precise the data is. 

```{r, echo = FALSE}
data_calibrated %>% 
  # look only at the standards
  filter(type == "standard" & is_standard) %>% 
  # check for each compound (+ other run groupings if appropriate)
  group_by(compound) %>% 
  # generate summary table comparing the 
  iso_generate_summary_table(
    raw_d13C = `d 13C/12C`, 
    true_d13C, 
    # this is the value predicated by the calibration
    d13C_pred = true_d13C_pred, 
    # and this is the estimated error based on the calibration
    d13C_error = true_d13C_pred_se
  )
```

### Visualization

Keep in mind that the calibration provides not just the predicted value of a peak but also the standard error based on the calibration and an indicator whether a predicted data point was in the calibration range or not. 

```{r "data_visualization", fig.width=12, fig.height=10}
data_calibrated %>% 
  iso_plot_data(
    x = Analysis, 
    y = true_d13C_pred, 
    y_error = true_d13C_pred_se,
    color = true_d13C_pred_in_range,
    shape = type,
    points = TRUE,
    lines = FALSE
  ) + 
  # additional features beyond the default plot
  facet_wrap(~compound, scales = "free_y") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  labs(color = "in calib. range")
```


# Combine with other data for final plots
## Remove bad chromtogram data
```{r}
final_data <- data_calibrated %>%
  #filter out bad chromatograms - all are the lower amount injection, which were followed by a larger volume - keeping the larger volume injection data
  filter(Analysis != "597") %>% # 122 depth, OG005
  filter(Analysis != "593") %>% # 121.105, OG047
  filter(Analysis != "612") %>% # 121.89, OG040
  filter(Analysis != "613") %>% # 120.205, OG042-2
  filter(Analysis != "609") %>% # 119.55, OG158
  filter(Analysis != "595") %>% # 122.9, OG043
  filter(type == "sample") %>%
  filter(true_d13C_pred_in_range == TRUE) %>%
  select(
      # sample info
      depth = `Depth`,  `compound`, Identifier.1 = `Identifier 1`, `ampl_sample_mean.mV`,  type, Analysis, 
      # peak info
       Ampl.44 = `Ampl 44`,  true_d13C_pred , true_d13C_pred_se, true_d13C_pred_in_range,  d13C_resid
    ) %>%
  mutate(
    prep = "adduct"
  )

vis <- final_data %>% 
  filter(compound %in% c("C29", "C27", "C31", "C33"))%>%
  iso_plot_data(
    x = depth, 
    y = true_d13C_pred, 
    y_error = true_d13C_pred_se,
    size = Ampl.44
  ) + 
  # additional features beyond the default plot
  facet_grid(~compound, scales = "free") + 
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  labs(y = "d13C") +
  scale_x_reverse() +
  coord_flip()
ggplotly(vis)

```

## Low res alkane data
```{r}
lowres <- read.csv("lowres_d13c.csv") %>% select( -X, -`measurement_info`, -`file_datetime`, -file_id) %>% mutate(prep = "non-adduct") 


sapply(final_data, class)
sapply(lowres, class)

final_data <- bind_rows(lowres, final_data)
```


# Bring in complimentary data (Jones et al., 2018)
## Inorganic geochem data

```{r}
cisotope <- read_excel(file.path("metadata", "Appendix_Table1_geochemistry.xlsx")) %>% 
  #rename columns
  select(depth = `Abs. depth (m)` , d13c_org = `Average δ13Corg (‰ VPDB)` )

final_data <- full_join(cisotope, final_data, by = "depth")
```

## Fit to floating timescale
```{r}

timescale <- read_excel(file.path("metadata", "Appendix_Table2_SH1_agemodel.xlsx")) %>% 
  rename(depth = `Depth`) %>%
  arrange(depth)

# extrapolation
extrapolated_timescale <-
  data_frame(
    depth = final_data$depth %>% unique()
  ) %>% 
  filter(!is.na(depth)) %>% 
  mutate(
    in_range = depth <= max(timescale$depth) & depth >= min(timescale$depth),
    Date.Ma = map_dbl(depth, function(d) {
      fit_data <- mutate(timescale, closest = abs(d - depth)) %>% arrange(closest)
      m <- lm(Date.Ma ~ depth, fit_data[1:10,])
      p <- predict(m, data.frame(depth = d), se.fit = TRUE)
      return(p$fit)
    }) ,
    in_range = depth <= max(timescale$depth) & depth >= min(timescale$depth),
    ka = map_dbl(depth, function(d) {
      fit_data <- mutate(timescale, closest = abs(d - depth)) %>% arrange(closest)
      m <- lm(ka ~ depth, fit_data[1:10,])
      p <- predict(m, data.frame(depth = d), se.fit = TRUE)
      return(p$fit)
    }))  
    
ggplot() + 
  aes(depth, Date.Ma) + 
  geom_line(data = timescale) +
  geom_point(data = extrapolated_timescale, mapping = aes(color = in_range))

alldata_w_time <- left_join(final_data, extrapolated_timescale, by = "depth")
```


#Compare d13C_DIC estimate with d13C_CaCO3 from Pueblo
## Add d13C_CaCO3 from planktonic forams from Keller et al. (2004) 
```{r}
# Used Jones et al. (2018) d13C org curve to correlate with org curve in Keller et al.(2004) for temporal correlation
CaCO3 <- data_frame(
Date.Ma = c(94.56738,  94.44, 94.38, 94.2, 93.94, 93.8, 93.57702), 
d13C = c(-0.2,  0.5, 2, 1.75, 2.5, 1, 1)
) 

# base plot
CaCO3 %>% 
  ggplot() +
  aes(x = Date.Ma, y = d13C) + 
  geom_line() +
  geom_point() +
  scale_x_reverse()+
  coord_flip()


# extrapolation
extrapolated_carbonate <-
  data_frame(
    Date.Ma = alldata_w_time$Date.Ma %>% unique()
  ) %>% 
  filter(!is.na(Date.Ma)) %>% 
  arrange(Date.Ma) %>%
  mutate(
    in_range = Date.Ma <= max(CaCO3$Date.Ma) & Date.Ma >= min(CaCO3$Date.Ma),
    d13C = map_dbl(Date.Ma, function(d) {
      fit_data <- mutate(CaCO3, closest = abs(d - Date.Ma)) %>% arrange(closest)
      m <- lm(d13C ~ Date.Ma, fit_data[1:3,])
      p <- predict(m, data.frame(Date.Ma = d), se.fit = TRUE)
      return(p$fit)
    }) )  

extrapolated_carbonate %>%
  ggplot() + 
  aes( Date.Ma, d13C) + 
  geom_line() +
  scale_x_reverse()+
  coord_flip()

alldata_w_time <- left_join(alldata_w_time, extrapolated_carbonate, by = "Date.Ma")

alldata_w_time %>%
  ggplot()+
  aes(Date.Ma, d13C) +
  geom_point() +
  geom_line() +
  scale_x_reverse()+
  coord_flip()
```


## Calculate / estimate eps_p
```{r}
alldata_w_time <- alldata_w_time %>%
    mutate(
         d13C_p_min = case_when(.$compound %in% c("C17", "C19", "C21")  ~ true_d13C_pred + 4,
                           .$compound == "phytane" ~ true_d13C_pred + 4.4 ) ,  
         d13C_p_max = case_when(.$compound %in% c("C17", "C19", "C21")  ~ true_d13C_pred + 8.4,
                           .$compound == "phytane" ~ true_d13C_pred + 6.4 ) ,
         d13C_p_avg = (d13C_p_max + d13C_p_min)/2, 
         
        eps_f = case_when(.$compound %in% c("C17", "C19", "C21")  ~ 24,
                           .$compound == "phytane" ~ 25 ) ,    # from Julio's unpublished text, 22 for alkanes (cyano-plank intermediate) and 25 for phytane; Hays 2001 says 16-22 for cyanos; Pagani 2002 says 25 for phytoplankton in general;  "Goericke et al. [ 1994] calculate the likely range of eps_f for marine phytoplankton to be 25-28" (from Brigidare et al., 1997). After Modelling (below), 24 for short chains gets us closest to phytane value (assuming 25 for phytane) - has implications for makeup of community, i.e., that cyanos are lower abundance than phytoplankton
         
         b = 200, # per Julio's unpublished text, can sequence from [118 to 262], or use constant 170. I think at least 170, potentially higher, given modelled need for high [PO4] advection to sustain OAE2
         
         Ko = (2.466 * 10^-2) # solubility constant, from Weiss 1974 (tables II and III); dep. on T and Salinity, range from ~2 to ~7 over all regimes; @ ~35 C, range is from 2.1 to 2.3 across all salinities. Choosing for 30ppt (34 or lower in Hay et al., 1996; 20-32 for shallow WIS in Petersen et al., 2016) and 32C (mid range T from O'Brient et al., 2017; from Blake Nose)
         
    )
```


## Calculate / estimate eps_p using CaCO3 rather than lipids
```{r}
alldata_w_time <- alldata_w_time %>%
    mutate(
        Tk = 32 + 273.15, #T from O'brien et al 2017 d18O planktonic forams from ODP site 1050 (Blake Nose - closest in lat and long to WIS), norm to Kelvin
         
         eps_b = 24.12 - (9866 / Tk), # temp dependent carbon isotope fractionation of CO2 - HCO3 ; from Julio's unpublished text
         
         d13C_aq_CO2 = d13C - 1 + eps_b, # from Julio's unpublished text and many others e.g., Van Bentum et al., 2012 , using CaCO3 d13C from Keller et al (above chunk)
        
         eps_p_carb = 1000 * ((d13C_aq_CO2 + 1000)/ (d13C_p_avg + 1000) -1), # from Freeman and Hayes 1992
        
         CO2_aq_conc_carb = b / (eps_f - eps_p_carb) , # from Brigidare et al. 1997
         
         pCO2_carb = CO2_aq_conc_carb / Ko  
        
        )
```

## Check calcultions immediately
```{r}

alldata_w_time %>% filter(compound == "C21") %>% select(compound, eps_b, eps_p_carb,  d13C_aq_CO2, pCO2_carb) %>% kable()
alldata_w_time %>% filter(compound == "phytane") %>% select(compound, eps_b, eps_p_carb,  d13C_aq_CO2, pCO2_carb) %>% kable()

```


## Model role of measured and fixed variables in pCO2 output
```{r}
data_try <- expand.grid(
b = seq(170, 262, by = 10), 
Ko = seq(2.2, 2.6, by = 0.2) * 10^-2,
eps_p = 15.5,
eps_f = seq(22, 25, by = 1)
) %>% as_data_frame() %>%
  mutate(
    CO2_aq = b / (eps_f - eps_p),
    pCO2 = CO2_aq / Ko
)

# base plot
data_try %>%
  iso_plot_data(x = b, y = pCO2, linetype = factor(Ko), color = factor(eps_f)) + 
  geom_line(size = 3) +
  theme(text = element_text(size=40)) 

  

```

# Final plots

## High res
```{r "", fig.width=12, fig.height=10}

alkane <- alldata_w_time %>% 
  filter(compound %in% c( "C27", "C29", "C31", "C33"), true_d13C_pred_in_range == TRUE)%>%
  filter(Identifier.1 != "OG005 AD")%>%
  ggplot(
    aes(
    x = Date.Ma, 
    y = true_d13C_pred, 
    y_error = true_d13C_pred_se ,
    color = compound
  ) )+ 
  geom_rect( 
    mapping = aes(
      xmin= -Inf, 
      xmax= 94.42,
      ymin= -Inf,
      ymax= Inf
    ),
    color = "grey",
    fill = "grey",
    alpha=0.5
    ) +
  geom_point(size =  6) +
  geom_line()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  labs(y = "d13C") +
  scale_x_reverse(limits = c(94.5, 94.34)) +
  coord_flip() +
  theme(text = element_text(size=40))
ggplotly(alkane)


alkanesh <- alldata_w_time %>% 
  filter(compound %in% c( "C19", "C21", "phytane"), true_d13C_pred_in_range == TRUE)%>%
  filter(Identifier.1 != "OG005 AD")%>%
  iso_plot_data(
    x = Date.Ma, 
    y = true_d13C_pred, 
    y_error = true_d13C_pred_se ,
    #shape = compound ,
    color = compound
  ) + 
  # additional features beyond the default plot
  #facet_grid(~compound, scales = "free") + 
  geom_point(size = 8) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  labs(y = "d13C") +
  scale_x_reverse(limits = c(94.5, 94.34)) +
  #scale_x_continuous(limits = c(94.6, 94.3))+
  #scale_y_continuous(limits = c(-33, -26)) +
  coord_flip() +
    theme(text = element_text(size=40))
ggplotly(alkanesh)


bulkorg <-  subset(alldata_w_time) %>%
  iso_plot_data(
    x = Date.Ma, 
    y = d13c_org
  ) + 
  # additional features beyond the default plot
  geom_point(size = 5) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  labs(y = "d13C") +
  scale_x_reverse(limits = c(94.5, 94.34)) +
   theme(panel.grid.minor.x = element_blank())+
  #scale_x_continuous(limits = c(-100, 120)) +
  #coord_cartesian(xlim=c(-20, 80))+
  coord_flip()+
    theme(text = element_text(size=40))
ggplot(bulkorg)

grid.arrange(bulkorg, alkane, ncol = 2)
```

## Low res (all)
```{r "", fig.width=12, fig.height=10}
alkaneall <- alldata_w_time %>% 
  filter(compound %in% c("pristane", "phytane", "C19", "C21"), true_d13C_pred_in_range == TRUE)%>%
  iso_plot_data(
    x = Date.Ma, 
    y = true_d13C_pred, 
    y_error = true_d13C_pred_se ,
    color = compound
  ) + 
  # additional features beyond the default plot
  #facet_grid(~compound, scales = "free", space = "free") + 
  geom_point(size = 5) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  labs(y = "d13C") +
  scale_x_reverse(limits = c(94.6,  93.6 )) +
  #scale_x_continuous()+
  scale_y_continuous() +
  coord_flip() +
   theme(text = element_text(size=40))

ggplotly(alkaneall)

bulkorgall <-  subset(alldata_w_time) %>%
  iso_plot_data(
    x = Date.Ma, 
    y = d13c_org
  ) + 
  # additional features beyond the default plot
  geom_point(size = 5) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  labs(y = "d13C") +
  scale_x_reverse(limits = c(94.6,  93.6 )) +
   theme(panel.grid.minor.x = element_blank())+
  coord_flip()+
   theme(text = element_text(size=40))
ggplotly(bulkorgall)

grid.arrange(bulkorgall, alkaneall, ncol = 2)
```

## Pool composition estimates from lipids
```{r}
pools <- alldata_w_time %>% 
   filter(Identifier.1 != "OG005_F1 alkanes_122m") %>% 
  filter(compound %in% c( "C31", "C33"))%>%
  ggplot()+
  facet_grid(~compound, scales = "free", space = "free") + 
  geom_point(aes(x = Date.Ma , y = d13C_pool_avg), size = 5) +
  #geom_errorbar(aes(x = ka , ymin = d13C_pool_min, ymax = d13C_pool_max)) +
  #geom_errorbarh(aes(xmin = age_dev, xmax = ka , y = d13C_pool_avg)) +
  #geom_line(aes(x = age_dev, y = d13C_pool_avg, color = "red")) +
  #geom_line(aes(x = ka, y = d13C_pool_max)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  labs(y = "Inorganic pool d13C") +
  scale_x_reverse(limits = c(94.35, 94.5)) +
  #scale_x_continuous()+
  scale_y_continuous() +
  #scale_x_continuous(limits = c(94.35, 94.5)) +
  #coord_cartesian(xlim=c(-20, 80))+
  coord_flip()+
  theme(text = element_text(size=40))
ggplotly(pools)
```

## Compare pool composition estimates from lipids  with CaCO3
```{r}
pool_DIC_lipid <- alldata_w_time %>% 
  filter(compound %in% c("phytane"))%>%
  ggplot()+
  facet_grid(~compound, scales = "free", space = "free") + 
  geom_point(aes(x = Date.Ma , y = d13C_pool_avg)) +
  #geom_errorbar(aes(x = ka , ymin = d13C_pool_min, ymax = d13C_pool_max)) +
  #geom_errorbarh(aes(xmin = age_dev, xmax = ka , y = d13C_pool_avg)) +
  #geom_line(aes(x = age_dev, y = d13C_pool_avg, color = "red")) +
  #geom_line(aes(x = ka, y = d13C_pool_max)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  labs(y = "Inorganic pool d13C from lipid") +
  scale_x_reverse() +
  #scale_x_continuous()+
  scale_y_continuous() +
  #scale_x_continuous(limits = c(-100, 120)) +
  #coord_cartesian(xlim=c(-20, 80))+
  coord_flip()

pool_DIC_carb <- alldata_w_time %>% 
  filter(compound %in% c( "phytane"))%>%
  ggplot()+
  facet_grid(~compound, scales = "free", space = "free") + 
  geom_point(aes(x = Date.Ma , y = d13C_aq_CO2)) +
  #geom_errorbar(aes(x = ka , ymin = d13C_pool_min, ymax = d13C_pool_max)) +
  #geom_errorbarh(aes(xmin = age_dev, xmax = ka , y = d13C_pool_avg)) +
  #geom_line(aes(x = age_dev, y = d13C_pool_avg, color = "red")) +
  #geom_line(aes(x = ka, y = d13C_pool_max)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  labs(y = "Inorganic pool d13C from CaCO3") +
  scale_x_reverse() +
  #scale_x_continuous()+
  scale_y_continuous() +
  #scale_x_continuous(limits = c(-100, 120)) +
  #coord_cartesian(xlim=c(-20, 80))+
  coord_flip()

grid.arrange(pool_DIC_carb, pool_DIC_lipid, ncol=2)
```

## Pool estimates with temporal offset estimates
```{r}
pools_alttime <- alldata_w_time %>% 
  filter(Identifier.1 != "OG005_F1 alkanes_122m") %>% #removing duplicate for 94.45 Ma (non-adducted)
  filter(compound %in% c( "C21",  "C33"))%>%
  ggplot()+
  aes(x = Date.Ma , y = d13C_pool_avg, color = "red") +
  facet_grid(~compound, scales = "free", space = "free") + 
  geom_line() +
  #geom_errorbar(aes(x = ka , ymin = d13C_pool_min, ymax = d13C_pool_max)) +
  #geom_errorbarh(aes(xmin = age_dev, xmax = ka , y = d13C_pool_avg)) +
  #geom_line(aes(x = ka, y = d13C_pool_min)) +
  geom_line(aes(x = age_dev, y = d13C_pool_avg, color = "blue")) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  labs(y = "Inorganic pool d13C") +
  scale_x_reverse(limits = c(94.6, 94.3)) +
  #scale_x_continuous()+
  scale_y_continuous() +
  #scale_x_continuous(limits = c(94.6, 94.55)) +
  #coord_cartesian(xlim=c(-20, 80))+
  coord_flip()

ggplotly(pools_alttime)

```


## CO2 calculations by date
```{r}
alldata_w_time %>% filter(compound == "phytane") %>% select(Date.Ma, eps_p_lipid, d13C_pool_avg, true_d13C_pred, CO2_aq_conc, pCO2) %>% kable() 
```

## eps_p
```{r}
#using my eps_p (lipd rather than caco3)
eps_p_lipid <- alldata_w_time %>% 
  filter(compound %in% c( "phytane" , "C19", "C21"), true_d13C_pred_in_range == TRUE)%>% 
  iso_plot_data(
    x = Date.Ma, 
    y = eps_p_lipid 
    #y_error = true_d13C_pred_se ,
    #shape = compound 
    #color = prep
  ) + 
  # additional features beyond the default plot
  facet_grid(~compound, scales = "free") + 
  geom_point() +
  #theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  labs(y = "eps_p_lipid") +
  scale_x_reverse() +
  #scale_x_continuous(limits = c(-100, 120))+
  #scale_y_continuous(limits = c(-33, -26)) +
  coord_flip() 

eps_p_carb <- alldata_w_time %>% 
  filter(compound %in% c("C19", "C21", "phytane"), true_d13C_pred_in_range == TRUE)%>% 
  iso_plot_data(
    x = Date.Ma, 
    y = eps_p_carb ,
    #y_error = true_d13C_pred_se ,
    #shape = compound 
    color = compound
  ) + 
  # additional features beyond the default plot
  #facet_grid(~compound, scales = "free") + 
  geom_point(size = 5) +
  #theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  labs(y = "eps_p") +
  scale_x_reverse(limits = c(94.6,  93.6 )) +
  #scale_x_continuous(limits = c(-100, 120))+
  #scale_y_continuous(limits = c(-33, -26)) +
  coord_flip()+
   theme(text = element_text(size=40))

grid.arrange(eps_p_lipid, eps_p_carb, ncol = 2)
```

## pCO2
```{r}
#using my eps_p (lipd rather than caco3)
pCO2 <- alldata_w_time %>% 
  filter(compound %in% c( "phytane" , "C21"), true_d13C_pred_in_range == TRUE)%>% 
  iso_plot_data(
    x = Date.Ma, 
    y = pCO2 
    #y_error = true_d13C_pred_se ,
    #shape = compound 
    #color = prep
  ) + 
  # additional features beyond the default plot
  facet_grid(~compound, scales = "free") + 
  geom_point() +
  #theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  labs(y = "pCO2 using lipid") +
  scale_x_reverse() +
  #scale_x_continuous(limits = c(-100, 120))+
  #scale_y_continuous(limits = c(-33, -26)) +
  coord_flip() 
ggplotly(pCO2)

pCO22carb <- alldata_w_time %>% 
  filter(compound %in% c("phytane", "C19", "C21"), true_d13C_pred_in_range == TRUE)%>% 
  filter(Identifier.1 != "OG005 AD") %>% #removing duplicate for 94.45 Ma (non-adducted)
  iso_plot_data(
    x = Date.Ma, 
    y = pCO2_carb, 
    #y_error = true_d13C_pred_se ,
    #shape = compound 
    color = compound
  ) + 
  # additional features beyond the default plot
  #facet_grid(~compound, scales = "free") + 
  geom_point(size = 8) +
  geom_line()+ 
  #theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  labs(y = "pCO2") +
  scale_x_reverse(limits = c(94.5, 94.34)) +
  #scale_x_continuous(limits = c(-100, 120))+
  #scale_y_continuous(limits = c(-33, -26)) +
  coord_flip() +
   theme(text = element_text(size=40))
ggplotly(pCO22carb)

pCO2carb <- alldata_w_time %>%
  filter(compound %in% c( "C19", "C21"), true_d13C_pred_in_range == TRUE)%>% 
  iso_plot_data(
    x = Date.Ma, 
    y = pCO2_carb ,
    #y_error = true_d13C_pred_se ,
    #shape = compound 
    color = compound
  ) + 
  # additional features beyond the default plot
  #facet_grid(~compound, scales = "free") + 
  geom_point(size = 5) +
  #geom_smooth()+ 
  #theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  labs(y = "pCO2 using CaCO3") +
  scale_x_reverse()+
  #scale_x_continuous(limits = c(-100, 120))+
  #scale_y_continuous(limits = c(-33, -26)) +
  coord_flip()+
   theme(text = element_text(size=20))
  ggplotly(pCO2carb)

alkane <- alldata_w_time %>% 
  filter(compound %in% c( "C31", "C33"), true_d13C_pred_in_range == TRUE, prep == "adduct")%>%
  ggplot()+
  aes(
    x = Date.Ma, 
    y = d13C_pool_avg, 
    #y_error = true_d13C_pred_se ,
    shape = compound 
    #color = prep
  ) + 
  # additional features beyond the default plot
  facet_grid(~compound, scales = "free") + 
  geom_smooth() +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  labs(y = "d13C") +
  scale_x_reverse(limits = c(94.5, 94.34)) +
  #scale_x_continuous(limits = c(94.6, 94.3))+
  #scale_y_continuous(limits = c(-33, -26)) +
  coord_flip() 


grid.arrange( pCO2, pCO2carb, ncol = 2)
```




