---
title: "20180712_GB_SH1_d13Call"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

```{r "cleanup", echo = FALSE, eval=FALSE}
# not run in knitted doc
rm(list = ls())
.rs.restartR()
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}

library(isoprocessor)
library(readxl)
library(tidyverse)
library(plotly)
library(knitr)
library(isoreader)
library(gridExtra)
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```


# Load data
```{r, echo = FALSE}
# load dataset
iso_files <- iso_read_continuous_flow("Garrett_Adducts.cf.rda")
```

## Vendor data table

```{r, echo = FALSE}
vdt <- iso_files %>% 
  # get relevant vendor data table columns and file info
  iso_get_vendor_data_table(
    select = c(starts_with("Amp"), everything()),
    include_file_info = c(ox = `Seed Oxidation`, everything())
  )
```

## Map peaks
```{r, echo=FALSE, eval=FALSE}
# file names
vdt$file_id %>% unique() %>%  paste(collapse = "\n") %>% message()
```

## Chromatograms
```{r, echo = FALSE, fig.width=10, fig.height=8, message=FALSE}
iso_files %>% 
  # fetch a few of the files of interst
  iso_filter_files(parse_number(Analysis) %in% c(603, 595)) %>% 
  # plot just mass 28
  iso_plot_continuous_flow_data(data = c(46)) %>% 
  # make interactive
  ggplotly(dynamicTicks = TRUE)
```

## Retrieve metadata

```{r, echo = FALSE}
metadata_samples <- read_excel(file.path("metadata", "20180416_GB_metadata.xlsx"), sheet = "files")
metadata_peak_maps <- read_excel(file.path("metadata","20180416_GB_metadata.xlsx"), sheet = "maps")
metadata_samples
```

## Map peaks

```{r, results='asis'}
# add metadata
data_w_metadata <- vdt %>% 
  iso_add_metadata(metadata_samples, match_by = c(`Identifier 1`)) 

# missing metadata
data_w_metadata %>% 
  iso_get_missing_metadata(select = c(Analysis, `Identifier 1`, map_id))

# map peaks
data_w_peaks_all <- data_w_metadata %>% 
  # only continue with records that have metadata and are flagged for processing
  iso_remove_missing_metadata() %>% 
  filter(process == "yes") %>% 
  # map peaks
  iso_map_peaks(metadata_peak_maps, file_id = file_id, rt = Rt, rt_start = Start, rt_end = End)

# problematic peaks
prob <- data_w_peaks_all %>% 
  iso_get_problematic_peaks(select = c(file_id, Analysis, compound, Start, Rt, End))

# focus on non problematic peaks only?
data_w_peaks <- data_w_peaks_all %>% iso_remove_problematic_peaks()
data_w_peaks
```

## Reference peaks

```{r "ref_variation", fig.width=8, fig.height=6}
data_w_peaks %>% 
  iso_plot_ref_peaks(
    is_ref_condition = is_ref_peak == "yes", 
    ratio = c(`R 13C/12C`, `R 18O/16O`),
    group_id = `Analysis`,
    within_group = TRUE
  ) 
```


# Data processing

## Focus on analytes 

Focus on the analytes and calculate a few summary parameters we want to use later.

```{r, echo = FALSE}
data_w_analyte_peaks <- 
  data_w_peaks %>% 
  # this is important so that the reference peaks are not caught up in the next set of calculations
  filter(is_ref_peak == "no") %>% 
  # for each analysis calculate averages across analysis
  group_by(file_id) %>% 
  mutate(
    ampl_sample_mean.mV = mean(`Ampl 44`), ampl_sample_sd.mV = sd(`Ampl 44`),
    area_sample_mean.Vs = mean(`Intensity 44`), area_sample_sd.Vs = sd(`Intensity 44`)
  )
```

## Evaluate Standards

### Known isotope values

Add known isotope values for standards.

```{r, echo = FALSE}
standards <- 
  read_excel(file.path("metadata", "GC-IRMS standards.xlsx")) %>% 
  mutate(type = "standard") %>% 
  mutate(compound = str_replace(compound, "-", ""))
standards

data_w_stds <-
  data_w_analyte_peaks %>% 
  iso_add_standards(standards, match_by = c(type, compound))  
data_w_stds
```

### Single analysis calibration (for QA)

Determine calibrations fits for individual standard analyses.

```{r, echo = FALSE}
stds_w_calibs <- data_w_stds %>%
  # focus on standard analyses
  filter(type == "standard") %>%  
  # prepare for calibration by defining the grouping column(s) 
  iso_prepare_for_calibration(group_by = c(Analysis)) %>% 
  # pull out additional ID columns from the nested data
  iso_unnest_data(
    select = c(`Analysis`, `Injection Volume` = `Identifier 2`)) %>% 
  # run calibration
  iso_generate_calibration(
    model = lm(`d 13C/12C` ~ true_d13C), 
    calibration = "d13C",
    is_standard = is_standard,
    # only consider the standards that actually have the 8 peaks
    min_n_datapoints = 8
  )

# check for problematic calibrations and make note of them in the main dataset
problematic_stds <- stds_w_calibs %>% iso_get_problematic_calibrations("d13C")
problematic_stds
data_w_stds <- data_w_stds %>% 
  mutate(type = ifelse(Analysis %in% problematic_stds$Analysis,
                       "excluded standard", type))

# move forward only with good calibrations
stds_w_calibs <- stds_w_calibs %>% iso_remove_problematic_calibrations("d13C")
```

### Calibration coefficients

Look at calibration parameters (coefficients and summary) as well ad data ranges in overview table.

```{r, echo = FALSE}
# look at coefficients and summary
stds_w_calibs %>% 
  # pull out a little more file information
  iso_unnest_data(select = c(`Seed Oxidation`)) %>% 
  # unnest calibration parameters
  iso_unnest_calibration_parameters(
    calibration = "d13C", 
    select_from_coefs = c(term, estimate, SE = std.error, signif),
    select_from_summary = c(fit_R2 = adj.r.squared, fit_RMSD = deviance, residual_df = df.residual)) %>% 
  arrange(term)

# look at ranges
stds_w_calibs %>% 
  iso_unnest_calibration_range(
    calibration = "d13C",
    select = c(var, min, max)
  ) %>% 
  arrange(var)
```

Visualize the calibration parameters

```{r "calibration_parameters", fig.width = 8, fig.height = 8}
# unnest some of the data for plotting
data_for_plots <- 
  stds_w_calibs %>%
  iso_unnest_data(
    c(Oxidation = `Seed Oxidation`, 
      datetime = file_datetime,
      `Mean Amplitude` = ampl_sample_mean.mV)
  ) 
data_for_plots


# parameters vs time
data_for_plots %>%
  iso_plot_calibration_parameters(
    calibration = "d13C", 
    x = datetime,
    color = `Injection Volume`,
    shape = Oxidation
  )

# parameter vs analysis
data_for_plots %>% 
  iso_plot_calibration_parameters(
    calibration = "d13C", 
    x = paste(Analysis),
    color = `Injection Volume`,
    shape = Oxidation
  )

# parameters vs amplitude
data_for_plots %>% 
  iso_plot_calibration_parameters(
    calibration = "d13C", 
    x = `Mean Amplitude`,
    color = `Injection Volume`
  ) 
```


```{r "interactive_calibration_parameters", fig.width = 8, fig.height = 8}
# turn the last plot into an interactive one
ggplotly(ggplot2::last_plot() + theme(legend.position = "none"))
```

### Inspect Residuals

Look at residuals to see goodness of fit for the single analysis calibrations.

```{r "standards_by_compound", fig.width = 8, fig.height = 8}
stds_w_calibs %>% 
  # pull out all data
  iso_unnest_data(select = everything()) %>% 
  filter(is_standard) %>% 
  # calculate deviation from means
  group_by(Analysis) %>% 
  mutate(
    `Var: residual d13C [permil]` = d13C_resid,
    `Var: area diff from mean [%]` = (`Intensity 44`/mean(`Intensity 44`) - 1) * 100,
    `Var: amplitude diff from mean [%]` = (`Ampl 44`/mean(`Ampl 44`) - 1) * 100
  ) %>%
  ungroup() %>% 
  # visualize
  iso_plot_data(
    x = compound,
    y = starts_with("Var"),
    group = Analysis,
    color = `Injection Volume`
  ) +
  # plot modifications
  labs(x = "")
```

```{r "interactive", fig.width = 8, fig.height = 8}
ggplotly()
```

## Gobal calibration

### Test different calibration models

Run global calibrations across all standards.

```{r, echo = FALSE}
#original
data_w_calibs <- data_w_stds %>% 
  # no grouping this time, everything is part of the global calibration
  iso_prepare_for_calibration() %>% 
  # run calibrations
  iso_generate_calibration(
    model = c(
      # simple model
      delta_only = lm(`d 13C/12C` ~ true_d13C),
      # multivariate with delta and amplitude (could also check Intensity 44)
      delta_and_ampl = lm(`d 13C/12C` ~ true_d13C + `Ampl 44`),
      # + the delta and amplitude cross term
      delta_cross_ampl = lm(`d 13C/12C` ~ true_d13C * `Ampl 44`),
      # multivariate with delta and the datetime (i.e. checking for temporal drift)
      delta_and_time = lm(`d 13C/12C` ~ true_d13C + file_datetime)
    ), 
    calibration = "d13C",
    # important to identify which analyses and which peaks within those should
    # be part of the calibration
    is_standard = type == "standard" & is_standard
  ) %>% 
  iso_remove_problematic_calibrations("d13C")
```


Visualize the calibration parameters for the different models. 

```{r "global_calibration_parameters", fig.width = 7, fig.height = 10}
# coefficient summary table
data_w_calibs %>% iso_unnest_calibration_parameters("d13C")

# plot to look at coefficients and summary
data_w_calibs %>% 
  iso_plot_calibration_parameters(
    calibration = "d13C",
    x = d13C_calib,
    color = d13C_calib,
    # include the significance level to gauge which model parameters matter
    shape = signif
  )
```


Garrett thinks delta and aplitude and delta cross amplitude look better, in fact

```{r "global_calibration_residuals", fig.width=8, fig.height=6}
data_w_calibs %>% 
  iso_unnest_data(select = everything()) %>% 
  rename(`Injection Volume` = `Identifier 2`) %>% 
  filter(type == "standard" & is_standard) %>% 
  # plot each standard analysis for each of the models to see the differences
  iso_plot_data(
    x = compound,
    y = d13C_resid,
    group = paste(Analysis, d13C_calib),
    color = d13C_calib
  ) 
```


### Apply calibration

Apply the `delta_and_ampl` calibration. 

```{r, echo = FALSE}
data_calibrated <- data_w_calibs %>% 
  # apply the delta_and_ampl calibration (could apply multiple and compare)
  filter(d13C_calib == "delta_and_ampl") %>% 
  iso_apply_calibration(
    predict = true_d13C,
    calibration = "d13C",
    calculate_error = TRUE
  ) %>% 
  # unnest the data
  iso_unnest_data()
data_calibrated
```

### Inspect results

The comparison of the predicted value in the standards along with the standard error in the prediction (the inherent error of the global calibration itself, which should be treated as the analytical error) can give a better sense of how precise the data is. 

```{r, echo = FALSE}
data_calibrated %>% 
  # look only at the standards
  filter(type == "standard" & is_standard) %>% 
  # check for each compound (+ other run groupings if appropriate)
  group_by(compound) %>% 
  # generate summary table comparing the 
  iso_generate_summary_table(
    raw_d13C = `d 13C/12C`, 
    true_d13C, 
    # this is the value predicated by the calibration
    d13C_pred = true_d13C_pred, 
    # and this is the estimated error based on the calibration
    d13C_error = true_d13C_pred_se
  )
```

### Visualization

Many different ways exist to visualize the data. Regardless of the plotting method it is helpful to keep in mind that the calibration provides not just the predicted value of a peak but also the standard error based on the calibration and an indicator whether a predicted data point was in the calibration range or not. 

```{r "data_visualization", fig.width=12, fig.height=10}
data_calibrated %>% 
  iso_plot_data(
    x = Analysis, 
    y = true_d13C_pred, 
    y_error = true_d13C_pred_se,
    color = true_d13C_pred_in_range,
    shape = type,
    points = TRUE,
    lines = FALSE
  ) + 
  # additional features beyond the default plot
  facet_wrap(~compound, scales = "free_y") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  labs(color = "in calib. range")
```


# Combine with other data for final plots
## Remove bad chromtogram data
```{r}
final_data <- data_calibrated %>%
  #filter out bad chromatograms - all are the lower amount injection, which were followed by a larger volume - keeping the larger volume injection data
  filter(Analysis != "597") %>% # 122 depth, OG005
  filter(Analysis != "593") %>% # 121.105, OG047
  filter(Analysis != "612") %>% # 121.89, OG040
  filter(Analysis != "613") %>% # 120.205, OG042-2
  filter(Analysis != "609") %>% # 119.55, OG158
  filter(Analysis != "595") %>% # 122.9, OG043
  filter(type == "sample") %>%
  filter(true_d13C_pred_in_range == TRUE) %>%
  select(
      # sample info
      depth = `Depth`,  `compound`, Identifier.1 = `Identifier 1`, `ampl_sample_mean.mV`,  type, Analysis, 
      # peak info
       Ampl.44 = `Ampl 44`,  true_d13C_pred , true_d13C_pred_se, true_d13C_pred_in_range,  d13C_resid
    ) %>%
  mutate(
    prep = "adduct"
  )

vis <- final_data %>% 
  filter(compound %in% c("C29", "C27", "C31", "C33"))%>%
  iso_plot_data(
    x = depth, 
    y = true_d13C_pred, 
    y_error = true_d13C_pred_se,
    size = Ampl.44
  ) + 
  # additional features beyond the default plot
  facet_grid(~compound, scales = "free") + 
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  labs(y = "d13C") +
  scale_x_reverse() +
  coord_flip()
ggplotly(vis)

```

## Low res alkane data
```{r}
lowres <- read.csv("lowres_d13c.csv") %>% select( -X, -`measurement_info`, -`file_datetime`, -file_id) %>% mutate(prep = "non-adduct") 


sapply(final_data, class)
sapply(lowres, class)

final_data <- bind_rows(lowres, final_data)
```

# Manipulate Alkane Data
## Correct for biosynth fractionation
```{r}
final_data <- final_data %>%
  filter(type == "sample") %>%
  select(-true_d13C, -d13C_resid) %>%
  filter(compound %in% c("phytane", "C17", "C19", "C21", "C27", "C29", "C31", "C33", "C35")) %>%
# net fractionation factor for inorganic -> lipid d13C (eps_net) calculated as = (inorganic -> biomass fractionation factor)  + (biomass -> lipid fractionation factor)

# minimum and max for marine algal biomass (first number in equation) from spread in data of algal biomass in Ohkouchi et al., 2015; OPTIONS FOR SHORTCHAIN: -4 per mil for phytane (Schouten et al., 1998; Sinninghe Damste et al., 2008), -8.4 per mil for cyano nC17 , -6.4 per mil for phytane (above takes into account frctionation in sulfur-bound and release per Sinninghe Damste?)(Sakata et al., 1997)
  
# lipid-leaf addition (second number in equation) by chain length from gymnosperm data in Diefendorf et al., 2011 [NOTE: average for gymnosperm is -2.5 per mil, that may be more appropriate]; minimum and maximum offset of leaf to atmosphere from C3 plant histogram in Ehleringer and Cerling, 2001 (in Taiz et al., 6th ed. Plant Physiology and Development)
  
  mutate(eps_net_min = case_when(.$compound %in% c("C17", "C19", "C21")  ~ (12+4) ,
                                 .$compound  == "phytane" ~ (12+4.4),
                                 .$compound  == "C27" ~ (22+2),  
                                 .$compound  == "C29" ~ (22+3),
                                 .$compound  == "C31" ~ (22+4),
                                 .$compound ==  "C33" ~ (22+3.5),
                                 .$compound ==  "C35" ~ (22+3)), 
        
#mutate(eps_net_min = case_when(.$compound %in% c("C17", "C19", "C21")  ~ (20+4) , #chose higher minimum fractionation for marine given high CO2 (not transport dominated) - makes pool value (and thus pCO2 estimates) way too high 
                                # .$compound  == "phytane" ~ (20+4),
                                # .$compound  == "C27" ~ (25+2),  # Chosing higher minimum for C3 plant photosynth bc carboxylation (Ci) is most likely limiting given high atm CO2 concentatiion (Ca) per O'leary 1998 "Carbon isotopes in photosynthesis" 
                                 #.$compound  == "C29" ~ (25+3),
                                 #.$compound  == "C31" ~ (25+4),
                                # .$compound ==  "C33" ~ (25+3.5),
                                # .$compound ==  "C35" ~ (25+3)), 
                
         eps_net_max = case_when(.$compound %in% c("C17", "C19", "C21")  ~ (25+8) , 
                                 .$compound  == "phytane" ~ (25+6.4),
                                 .$compound  == "C27" ~ (32+2),  
                                 .$compound  == "C29" ~ (32+3),
                                 .$compound  == "C31" ~ (32+4), 
                                 .$compound ==  "C33" ~ (32+3.5),
                                 .$compound ==  "C35" ~ (32+3)),  
         
         d13C_pool_min = true_d13C_pred + eps_net_min,
         
         d13C_pool_max = true_d13C_pred + eps_net_max,
         
         d13C_pool_avg = true_d13C_pred + ((eps_net_max + eps_net_min) / 2) )
```


## Calculate / estimate eps_p
```{r}
final_data <- final_data %>%
  #filter(compound %in% c("phytane","C17", "C19", "C21")) %>% #this will apply this eps_p to all chain lengths, but we'd only plot using the short chains. alternatively can "case_when" like for eps_f for all
    mutate(
         #d13C_CaCO3 = 2, # from Jones et al. avg of pre-excursion values (before diagenetic test shows high diagenesis); range from 0 to 4 in full 30m record
         
         #Tk = 35 + 273.15, #T from O'brien et al 2017 d18O planktonic forams from ODP site 1050 (Blake Nose - closest in lat and long to WIS), norm to Kelvin
         
         #eps_b = (24.12 - 9866) / Tk, # temp dependent carbon isotope fractionation of CO2 - HCO3 ; from Julio's unpublished text
         
         #d13C_aq_CO2 = d13C_CaCO3 - 1 + eps_b, # from Julio's unpublished text 
      
         #eps_p = 1000 * ((d13C_aq_CO2 + 1000)/ (d13C_p + 1000) -1), # from Freeman and Hayes 1992
         
         d13C_p = case_when(.$compound %in% c("C17", "C19", "C21")  ~ true_d13C_pred + 8.4,
                           .$compound == "phytane" ~ true_d13C_pred + 6.4 ) ,  # from Julio's unpublished text, original from Sakata et al., 1997 (6.4 in Sakata...)
         
         eps_p_lipid = 1000 * ((d13C_pool_avg + 1000)/ (d13C_p + 1000) -1), # from Freeman and Hayes 1992, but using avg estimated d13C of DIC from lipid rather than calc 
         
         eps_f = case_when(.$compound %in% c("C17", "C19", "C21")  ~ 24,
                           .$compound == "phytane" ~ 25 ) ,    # from Julio's unpublished text, 22 for alkanes (cyano-plank intermediate) and 25 for phytane; Pagani 2002 says 25 for phytoplankton in general;  "Goericke et al. [ 1994] calculate the likely range of eps_f for marine phytoplankton to be 25-28" (from Brigidare et al., 1997)
         
         b = 170, # per Julio's unpublished text, can sequence from [118 to 262], or use constant 170
         
         CO2_aq_conc = b / (eps_f - eps_p_lipid) , # from Brigidare et al. 1997
         
         Ko = (2.418 * 10^-2), # solubility constant, from Weiss 1974 (tables II and III); dep. on T and Salinity, range from ~2 to ~7 over all regimes; @ ~35 C, range is from 2.1 to 2.3 across all salinities. Choosing for 34ppt (Hay et al., 1996) and 32C (mid range T from O'Brient et al., 2017)
         
         pCO2 = CO2_aq_conc / Ko ) # from Julio's unpublished text , and Sarmiento and Gruber, Ocean Biogeochemical Dynamics, section 8.3 (p.327)
         
         #log_CO2_aq_conc = (eps_p_lipid / (-17 + 2.2)) - 3.4, # from Hayes et al., 1989 (eqn. 6), based on McCabe 1985
         
         #CO2_aq_conc = ((eps_p_lipid * 1.835) - 13.5)   # from Freeman and Hayes 1992, rsquared = 0.69, in micromol/kg
        
```

## Check calcultions immediately
```{r}

final_data %>% filter(compound == "C21") %>% select( compound, eps_p_lipid, d13C_pool_avg, true_d13C_pred, CO2_aq_conc, pCO2) %>% kable()

```

## Model role of measured and fixed variables in pCO2 output
```{r}
data_try <- expand.grid(
b = seq(170, 262, by = 10), 
Ko = seq(2.1, 2.9, by = 0.2) * 10^-2,
eps_p = seq(16.5, 17.5, by = 1)
) %>% as_data_frame() %>%
  mutate(
    eps_f = case_when(eps_p ==16.5 ~ 24,
                      eps_p == 17.5 ~ 25),
    compound = case_when (eps_p == 16.5 ~ "C17,19,21",
                          eps_p == 17.5 ~ "phytane"),
    CO2_aq = b / (eps_f - eps_p),
    pCO2 = CO2_aq / Ko
)


# base plot
base_plot <- 
  ggplot() +
  aes(x = b, y = pCO2, linetype = factor(Ko)) +
  geom_line() 

  
# plotting data 1
base_plot %+% data_try
```


## Model potential values from CaCO3 d13C values
```{r}
# d13C_p for C21 is ~-17 to -23

data_carb <- expand.grid(
d13C_l = -20, 
factorp = seq(6, 8, by = 2),
TC = 35,
d13C_CaCO3 = seq(0, 2, by = 1) ,
eps_f = 24
) %>% as_data_frame() %>%
  mutate(
    d13C_p = d13C_l + factorp, 
    Tk = TC + 273.15,
    eps_b = (24.12 - 9866) / Tk,
    d13C_aq_CO2 = d13C_CaCO3 - 1 + eps_b,
    eps_p = 1000 * ((d13C_aq_CO2 + 1000)/ (d13C_p + 1000) -1),
    CO2_aq = 170 / (eps_f - eps_p),
    pCO2 = CO2_aq / (2.4 * 10^-2)
)


# base plot
base_plot <- 
  ggplot() +
  aes(x = d13C_CaCO3, y = pCO2, color = factor(factorp)) +
  geom_line() 

  
# plotting data 1
base_plot %+% data_carb
```


# Bring in complimentary data (Jones et al., 2018)
## Inorganic geochem data

```{r}
cisotope <- read_excel(file.path("metadata", "Appendix_Table1_geochemistry.xlsx")) %>% 
  #rename columns
  select(depth = `Abs. depth (m)` , d13c_org = `Average δ13Corg (‰ VPDB)` )
  #remove columns not of interest
  #select(-`stdev δ13Corg`, -`δ13Ccarb stdev`, -`d18O-avg`, -`d18O stdev`, -`∆13C`, -`d13c_carb`) 

final_data <- full_join(cisotope, final_data, by = "depth")
```

## Fit to floating timescale
```{r}

timescale <- read_excel(file.path("metadata", "Appendix_Table2_SH1_agemodel.xlsx")) %>% 
  rename(depth = `Depth`) %>%
  arrange(depth)

# extrapolation
extrapolated_timescale <-
  data_frame(
    depth = final_data$depth %>% unique()
  ) %>% 
  filter(!is.na(depth)) %>% 
  mutate(
    in_range = depth <= max(timescale$depth) & depth >= min(timescale$depth),
    Date.Ma = map_dbl(depth, function(d) {
      fit_data <- mutate(timescale, closest = abs(d - depth)) %>% arrange(closest)
      m <- lm(Date.Ma ~ depth, fit_data[1:10,])
      p <- predict(m, data.frame(depth = d), se.fit = TRUE)
      return(p$fit)
    }) ,
    in_range = depth <= max(timescale$depth) & depth >= min(timescale$depth),
    ka = map_dbl(depth, function(d) {
      fit_data <- mutate(timescale, closest = abs(d - depth)) %>% arrange(closest)
      m <- lm(ka ~ depth, fit_data[1:10,])
      p <- predict(m, data.frame(depth = d), se.fit = TRUE)
      return(p$fit)
    }))  
    
ggplot() + 
  aes(depth, Date.Ma) + 
  geom_line(data = timescale) +
  geom_point(data = extrapolated_timescale, mapping = aes(color = in_range))

alldata_w_time <- left_join(final_data, extrapolated_timescale, by = "depth")
```

# Estimate error for deposition
```{r}
alldata_w_time <-  alldata_w_time %>%
  mutate(age_range = case_when(.$compound %in% c("C17", "C19", "C21", "phytane")  ~ (-0.00001) , 
                                 .$compound %in% c("C27", "C29", "C31", "C33", "C35") ~ (-.01)), #in MA (100 yrs, 10000 yrs)
         age_dev = Date.Ma + age_range 
   )
```


# Final plots

## High res
```{r "", fig.width=12, fig.height=10}

alkane <- alldata_w_time %>% 
  filter(compound %in% c( "C29", "C31", "C33"), true_d13C_pred_in_range == TRUE, prep == "adduct")%>%
  iso_plot_data(
    x = Date.Ma, 
    y = true_d13C_pred, 
    #y_error = true_d13C_pred_se ,
    shape = compound 
    #color = prep
  ) + 
  # additional features beyond the default plot
  facet_grid(~compound, scales = "free") + 
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  labs(y = "d13C") +
  scale_x_reverse(limits = c(94.5, 94.34)) +
  #scale_x_continuous(limits = c(94.6, 94.3))+
  #scale_y_continuous(limits = c(-33, -26)) +
  coord_flip() 
ggplotly(alkane)

bulkorg <-  subset(alldata_w_time, ka > -100 & ka < 120) %>%
  iso_plot_data(
    x = ka, 
    y = d13c_org
  ) + 
  # additional features beyond the default plot
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  labs(y = "d13C") +
  #scale_y_reverse() +
  scale_x_continuous(limits = c(-100, 120)) +
  #coord_cartesian(xlim=c(-20, 80))+
  coord_flip()


grid.arrange(bulkorg, alkane, ncol = 2)
```

## Low res (all)
```{r "", fig.width=12, fig.height=10}
alkaneall <- alldata_w_time %>% 
  filter(compound %in% c("pristane", "phytane", "C19", "C21", "C31", "C33"), true_d13C_pred_in_range == TRUE)%>%
  iso_plot_data(
    x = Date.Ma, 
    y = true_d13C_pred, 
    y_error = true_d13C_pred_se ,
    color = prep
  ) + 
  # additional features beyond the default plot
  facet_grid(~compound, scales = "free", space = "free") + 
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  labs(y = "d13C") +
  scale_x_reverse() +
  #scale_x_continuous()+
  scale_y_continuous() +
  coord_flip() 

ggplotly(alkaneall)

bulkorgall <-  subset(alldata_w_time) %>%
  iso_plot_data(
    x = ka, 
    y = d13c_org
  ) + 
  # additional features beyond the default plot
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  labs(y = "d13C") +
  #scale_y_reverse() +
  #scale_x_continuous(limits = c(-100, 120)) +
  #coord_cartesian(xlim=c(-20, 80))+
  coord_flip()


grid.arrange(bulkorgall, alkaneall, ncol = 2)
```

## Pool composition estimates
```{r}
pools <- alldata_w_time %>% 
  filter(compound %in% c("C19", "C21", "C31", "C33"))%>%
  ggplot()+
  facet_grid(~compound, scales = "free", space = "free") + 
  geom_point(aes(x = Date.Ma , y = d13C_pool_avg)) +
  #geom_errorbar(aes(x = ka , ymin = d13C_pool_min, ymax = d13C_pool_max)) +
  #geom_errorbarh(aes(xmin = age_dev, xmax = ka , y = d13C_pool_avg)) +
  #geom_line(aes(x = age_dev, y = d13C_pool_avg, color = "red")) +
  #geom_line(aes(x = ka, y = d13C_pool_max)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  labs(y = "Inorganic pool d13C") +
  scale_x_reverse() +
  #scale_x_continuous()+
  scale_y_continuous() +
  #scale_x_continuous(limits = c(-100, 120)) +
  #coord_cartesian(xlim=c(-20, 80))+
  coord_flip()
ggplotly(pools)
```


## Pool estimates with temporal offset estimates
```{r}
pools_alttime <- alldata_w_time %>% 
  filter(Identifier.1 != "OG005_F1 alkanes_122m") %>% #removing duplicate for 94.45 Ma (non-adducted)
  filter(compound %in% c( "C19",  "C33"))%>%
  ggplot()+
  aes(x = Date.Ma , y = d13C_pool_avg, color = "red") +
  facet_grid(~compound, scales = "free", space = "free") + 
  geom_line() +
  #geom_errorbar(aes(x = ka , ymin = d13C_pool_min, ymax = d13C_pool_max)) +
  #geom_errorbarh(aes(xmin = age_dev, xmax = ka , y = d13C_pool_avg)) +
  #geom_line(aes(x = ka, y = d13C_pool_min)) +
  geom_line(aes(x = age_dev, y = d13C_pool_avg, color = "blue")) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  labs(y = "Inorganic pool d13C") +
  scale_x_reverse(limits = c(94.6, 94.3)) +
  #scale_x_continuous()+
  scale_y_continuous() +
  #scale_x_continuous(limits = c(94.6, 94.55)) +
  #coord_cartesian(xlim=c(-20, 80))+
  coord_flip()

ggplotly(pools_alttime)

```


## CO2 calculations by date
```{r}
alldata_w_time %>% filter(compound == "phytane") %>% select(Date.Ma, eps_p_lipid, d13C_pool_avg, true_d13C_pred, CO2_aq_conc, pCO2) %>% kable() 
```

## eps_p
```{r}
#using my eps_p (lipd rather than caco3)
eps_p <- final_data %>% 
  filter(compound %in% c( "phytane" , "C19", "C21"), true_d13C_pred_in_range == TRUE)%>% 
  iso_plot_data(
    x = depth, 
    y = eps_p_lipid 
    #y_error = true_d13C_pred_se ,
    #shape = compound 
    #color = prep
  ) + 
  # additional features beyond the default plot
  facet_grid(~compound, scales = "free") + 
  geom_point() +
  #theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  labs(y = "eps_p") +
  #scale_x_reverse() +
  #scale_x_continuous(limits = c(-100, 120))+
  #scale_y_continuous(limits = c(-33, -26)) +
  coord_flip() 
ggplotly(eps_p)
```

## pCO2
```{r}
#using my eps_p (lipd rather than caco3)
pCO2 <- alldata_w_time %>% 
  filter(compound %in% c( "phytane" , "C19", "C21"), true_d13C_pred_in_range == TRUE)%>% 
  iso_plot_data(
    x = Date.Ma, 
    y = pCO2 
    #y_error = true_d13C_pred_se ,
    #shape = compound 
    #color = prep
  ) + 
  # additional features beyond the default plot
  facet_grid(~compound, scales = "free") + 
  geom_point() +
  #theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  labs(y = "pCO2") +
  scale_x_reverse() +
  #scale_x_continuous(limits = c(-100, 120))+
  #scale_y_continuous(limits = c(-33, -26)) +
  coord_flip() 
ggplotly(pCO2)

pCO22 <- alldata_w_time %>% 
  filter(compound %in% c( "C21"), true_d13C_pred_in_range == TRUE)%>% 
  ggplot()+
  aes(
    x = Date.Ma, 
    y = pCO2 
    #y_error = true_d13C_pred_se ,
    #shape = compound 
    #color = prep
  ) + 
  # additional features beyond the default plot
  facet_grid(~compound, scales = "free") + 
  geom_smooth() +
  geom_point() +
  #theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  labs(y = "pCO2") +
  scale_x_reverse(limits = c(94.5, 94.34)) +
  #scale_x_continuous(limits = c(-100, 120))+
  #scale_y_continuous(limits = c(-33, -26)) +
  coord_flip() 


alkane <- alldata_w_time %>% 
  filter(compound %in% c( "C31", "C33"), true_d13C_pred_in_range == TRUE, prep == "adduct")%>%
  ggplot()+
  aes(
    x = Date.Ma, 
    y = d13C_pool_avg, 
    #y_error = true_d13C_pred_se ,
    shape = compound 
    #color = prep
  ) + 
  # additional features beyond the default plot
  facet_grid(~compound, scales = "free") + 
  geom_smooth() +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  labs(y = "d13C") +
  scale_x_reverse(limits = c(94.5, 94.34)) +
  #scale_x_continuous(limits = c(94.6, 94.3))+
  #scale_y_continuous(limits = c(-33, -26)) +
  coord_flip() 

ggplotly(pCO2)
ggplotly(alkane)
grid.arrange( alkane, pCO22, ncol = 2)
```


# Statistical Analyses - a failed attempt
```{r}
# grouped time t-test
alldata_w_time %>% group_by(compound) %>% t.test(true_d13C_pred[ka = -100:-14], true_d13C_pred[ka = -18:20], true_d13C_pred[ka = -18:50], true_d13C_pred[ka = 20:50])

# by slope
lmdata <- lm(true_d13C_pred ~ ka, data=alldata_w_time)
print(lmdata)
```

```{r}
# ANOVA
##Ho = no change in true_d13C_pred thru time ; Ha = true_d13C_pred changes thru time
alldata_w_time %>%
  group_by(ka) %>%
  summarise(
    count = n(),
    mean = mean(true_d13C_pred, na.rm = TRUE),
    sd = sd(true_d13C_pred, na.rm = TRUE)
  )

# Compute the analysis of variance
res.aov <- aov(true_d13C_pred ~ ka, data = alldata_w_time)
# Summary of the analysis
summary(res.aov)

# Alternative ANOVA
install.packages("nlme")
install.packages("car")

library(nlme)

model = lme(true_d13C_pred ~ compound + ka + compound*ka, 
            random = ~1|compound,
            correlation = corAR1(form = ~ ka | compound,
                                 value = 0.4287),
            data=alldata_w_time,
            method="REML")

library(car)

Anova(model)

# check randomness
model.fixed = gls(true_d13C_pred ~ compound + ka + compound*ka,
                  data=alldata_w_time,
                  method="REML")

anova(model,
      model.fixed)


#correlation = Structure(form  = ~ time | subjvar)


library(rcompanion)

Sum = groupwiseMean(true_d13C_pred ~ compound + ka,
                    data   = alldata_w_time,
                    conf   = 0.95,
                    digits = 3,
                    traditional = FALSE,
                    percentile  = TRUE)

Sum

pd = position_dodge(.2)

ggplot(Sum, aes(x =    ka,
                y =    Mean)) +
    geom_errorbar(aes(ymin=Percentile.lower,
                      ymax=Percentile.upper),
                   width=.2, size=0.7, position=pd) +
    geom_point(shape=15, size=4, position=pd) +
    theme_bw() +
    theme(axis.title = element_text(face = "bold")) +
    ylab("Mean d13C_pred")

```

