---
title: "d13c_alkane_model"
author: "Garrett Boudinot"
date: "June 29, 2018"
output: html_document
---

```{r "load packages", message=FALSE, warning=FALSE, echo=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(tibble)
rm(list = ls())
knitr::opts_knit$set(root.dir = ".") 
```


#Setting up our model


#### Here's the equations in full:

 - $\delta^{13}C_DIC = (\delta^{13}C_CO2 * J_{CO2}) -((J_{bury} * \delta^{13}C_bury) + (J_{svol} * \delta^{13}C_svol))
 - $\delta^{13}C_DIC = ((\delta^{13}C_CO2 * F_{bury}) - \delta^{13}C_bury + (\delta^{13}C_svol * r)) * J_{bury}

 
With parameters:

 - $r = J_{svol}/J_{bury}$
 - $f_{bury} = J_{CO2}/J_{bury}$
 
And individual fluxes: 

 - $k = ((1/F_{bury}) * J_{svol})/(J_{CO2}*r)
 - $J_{bury} = J_{svol}/ (((1/F_{bury}) * J_{svol}) / (J_{CO2} * k))$
 - $J_{CO2} = (((1/F_{bury}) * J_{svol}) / (r * k))$
 - $J_{svol} = ((J_{CO2} * r * k) / f_{bury})$
 
 
#### Default values (data!)

 - $f_{bury} = seq(0.2, 1, by = 0.2)$
 - $r = seq(0.2, 1, by = 0.2)$
 - $\delta^{13}C_DIC = -10$
 - $\delta^{13}C_CO2 = -2$
 
 
 
 
 
 
 
```{r}
# constants

delta_13C_DIC = -10
delta_13C_CO2 = -2
delta_13C_svol = -1


# grid of variables
data1 <- expand.grid(
  f_bury = seq(0.2, 1, by = 0.2) ,
  r = seq(0.2, 1, by = 0.2),
  delta_13C_bury = seq(-30, -25, by = 1)
) %>% as_data_frame() %>% 
  mutate(
    J_bury = ((delta_13C_CO2 * f_bury) - delta_13C_bury + (delta_13C_svol * r)) / delta_13C_DIC
  )

# show the data
data1 %>% head() %>% knitr::kable()
```


#### Visualize

```{r}
# base plot
base_plot <- 
  ggplot() +
  aes(x = delta_13C_bury, y = J_bury, color = factor(r)) +
  geom_line() 

  
# plotting data 1
base_plot %+% data1
```
