---
title: "20180226_C_A6"
author: "Garrett Boudinot"
date: "`r format(Sys.Date(), '%d %b %Y')`"
output:
  html_document: 
    df_print: paged
    toc: yes
    toc_float: true
    toc_depth: 3
    code_folding: show
editor_options:
  chunk_output_type: inline
---

# Normal Start Functions

```{r "cleanup", echo = FALSE, eval=FALSE}
# not run in knitted doc
rm(list = ls())
.rs.restartR()
```

```{r setup, echo = FALSE, message=FALSE, warning=FALSE}
# load libraries
library(tidyverse) 
library(plotly)
library(isoreader)
library(dplyr)
library(ggplot2)
library(knitr)
library(readxl)
library(tidyr)
library(isoprocessorCUB)
source("functions.R")
# set global chunk options for figure output
figure_prefix <- "PREFIX-"
knitr::opts_chunk$set(
  dev=c("png", "pdf"), dev.args=list(pdf = list(encoding="WinAnsi", useDingbats=FALSE)),
  fig.keep="all", fig.path=file.path("figures", figure_prefix))
```

# Load isofiles

```{r, message=FALSE}
isofiles <-
  file.path(
    "Results",
    c(
      "180223_alkanes_C_GB"
    )
  ) %>%
  iso_read_continuous_flow()
```

# Data

## File Info

```{r}
# all file info
isofiles %>% iso_get_file_info()
```

## Vendor data table

```{r}
isofiles %>% iso_get_vendor_data_table()
```

## Chroms

```{r, fig.width=10, fig.height=8, message=FALSE}
isofiles %>% 
  # fetch a few of the files of interst
  iso_filter_files(parse_number(Analysis) %in% c(502, 504, 506)) %>% 
  # plot just mass 28
  iso_plot_continuous_flow_data(data = c(46)) %>% 
  # make interactive
  ggplotly(dynamicTicks = TRUE)
```

# Analysis

## Select relevant data

```{r}
data_table <-
  isofiles %>%
  # filter the files you want to use 
  # --> exclude CO2 zeros
  iso_filter_files(!str_detect(`Identifier 1`, "CO2 zero")) %>% 
  # --> select all analyses > 490
  iso_filter_files(parse_number(Analysis) > 490) %>% 
  # get the vendor data table and file info
  iso_get_vendor_data_table(
    select = c(
      # peak info
      Nr., Start, Rt, End,
      # amplitudes and intensities
      `Ampl 44`, `Ampl 46`, `Intensity All`,
      # ratios and deltas
      `R 46CO2/44CO2`, `d 13C/12C`
    ),
    include_file_info = c(
      file_datetime, Analysis, `Identifier 1`, `Identifier 2`, `GC Method`, `Seed Oxidation`
    )
  ) %>% 
  # rename some columns to be easier to work with
  rename(
    Ampl44=`Ampl 44`, Ampl46=`Ampl 46`, # amplitudes
    #Area28=`rIntensity 28`, Area29=`rIntensity 29`, # areas
    Intensity=`Intensity All`, #peak intensities
    R46C = `R 46CO2/44CO2`, d13C = `d 13C/12C`, # ratio and delta values
    rt = `Rt`,
    rt_start = `Start`,
    rt_end = `End`
    #File = `file_id`
  )
data_table
```

## Map peaks
```{r, echo=FALSE, eval=FALSE}
# file names
data_table$file_id %>% unique() %>%  paste(collapse = "\n") %>% message()
```

```{r}
### CHANGE MAPPING FILE NAME ###
metadata_samples <- read_excel(file.path("metadata", "20180226_GB_metadata.xlsx"), sheet = "files")

metadata_peak_maps <- read_excel(file.path("metadata","20180226_GB_metadata.xlsx"), sheet = "maps")

metadata_samples

#data_table_mapped <- map_peaks(data_table, metadata_file = file.path("metadata", "20180226_GB_metadata.xlsx"), quiet = FALSE)
```
```{r}
data_table_with_peaks <- data_table %>%
    iso_add_metadata(metadata_samples, match_by = c(`Identifier 1`, Analysis)) %>%
    iso_map_peaks(metadata_peak_maps) 

# missing and unidentified peaks
data_table_with_peaks %>% filter(!is_identified | is_missing)

# confirmed peaks
data_table_with_peaks <- data_table_with_peaks %>%
    filter(is_identified, !is_missing, !is_ambiguous)

data_table_with_peaks
```



## Check stability of reference peaks

The second one is the one defined to be 0 permil (so will always be), the rest is relative to that peak.

```{r, message=FALSE, fig.width = 10, fig.height = 8}
p <- data_table_with_peaks %>% 
  filter(is_ref_peak == "yes") %>% 
  ggplot() + 
  aes(Nr., d13C, color = file_id) +
  geom_line() + 
  theme_bw()
ggplotly(p)
```

## Add standard values

```{r}
standards <- read_excel(file.path("metadata", "gc_irms_indiana_A6.xlsx"))
data_w_stds <- data_table_with_peaks %>% 
  filter(type == "standard", is_ref_peak == "no") %>% 
  left_join(standards, by = "compound") %>% 
  mutate(is_std = !is.na(true.d13C) | !is.na(true.d2H))
```

## Evaluate standards

### Visualize

```{r, fig.width = 9, fig.height=6}
v <- data_w_stds %>% 
  ggplot() +
  aes(x = true.d13C, y = d13C, color = file_id) + 
  geom_smooth(method = "lm", se = FALSE, alpha = 0.5) +
  geom_point() +
  theme_bw() +
  theme(legend.position = "none") 
ggplotly(v)
```


### Overview

```{r}
std_corrections <- 
  data_w_stds %>% 
  group_by(file_id) %>% 
  do({
    m <- lm(d13C ~ true.d13C, data = .)
    data_frame(
      offset = coefficients(m)[1],
      slope = coefficients(m)[2],
      delta_ref_CO = -offset/slope,
      max_residual = max(summary(m)$residuals),
      rss = sum(summary(m)$residuals^2),
      r2 = summary(m)$r.squared
    )
  }) 
std_corrections %>% knitr::kable(d = 3)
```

## Intensities and delta values of the analyte peaks

Legend explanation:
 - color: TRUE/black fill = new runs, FALSE/gray fill = old runs
 - shape: 0/circle = no seed oxidation, 1/square = seed oxidation
 - fill: Identifier 2

```{r, message=FALSE, fig.width = 10, fig.height = 8}
# delta vs amplitude
p <- data_w_stds %>% 
  #filter(compound == "caffeine") %>% 
  ggplot() + 
  aes(Ampl44, d13C, fill = `Identifier 2`, group = Analysis) +
  geom_point() + 
  theme_bw() +
  scale_shape_manual(values = c(21:25, 3:4)) +
  scale_color_manual(values = c("gray", "black")) 
ggplotly(p, dynamicTicks = TRUE)

# delta vs time
p2 <- p %+% aes(x=file_datetime, size=Ampl44) +
  scale_x_datetime("", date_breaks = "12 hours", date_labels = "%d %b %H:%M")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
ggplotly(p2, dynamicTicks = TRUE)

# amplitude vs time
p3 <- p2 %+% aes(y = Ampl44, size = d13C)
ggplotly(p3, dynamicTicks = TRUE)
```


# Numerical Assessment

```{r}
# overall
data_w_stds %>% 
  filter(compound == "C21") %>% 
  # group into old and new runs
  # summarize information
  summarize(
    n_analyses = n(),
    d13C_mean = mean(d13C),
    d13C_sd = sd(d13C),
    Ampl44_mean = mean(Ampl44),
    Ampl44_sd = sd(Ampl44)
  )

# different injection volumes
assessment <-
  data_w_stds %>% 
  filter(compound == "C21") %>%
  # group into injection volume and new vs. old runs
  group_by(`Identifier 2`) %>%
  # summarize information
  summarize(
   n_analyses = n(),
    d13C_mean = mean(d13C),
    d13C_sd = sd(d13C),
    Ampl44_mean = mean(Ampl44),
    Ampl44_sd = sd(Ampl44)
  ) %>% ungroup() %>% 
  # sort to make it easy to see
  arrange(`Identifier 2`)

assessment
```

## Visualization

```{r, fig.width=10, fig.height=7}
p <- assessment %>% 
  ggplot() + 
  aes(Ampl44_mean, d13C_mean, fill = `Identifier 2`, text = n_analyses) + 
  geom_errorbar(map = aes(ymin = d13C_mean - d13C_sd, ymax = d13C_mean + d13C_sd), show.legend = FALSE) +
  geom_errorbarh(map = aes(xmin = Ampl44_mean - Ampl44_sd, xmax = Ampl44_mean + Ampl44_sd), show.legend = FALSE) +
  geom_point(shape = 21, size = 4) + 
  theme_bw() +
  scale_color_manual(values = c("gray", "black"))
ggplotly(p, dynamicTicks = TRUE)
```
