---
title: "processing Nadia's MAT253 standard runs"
author: "Garrett"
date: "7/20/2017"
output: html_document
---



```{r "cleanup", echo = FALSE, eval=FALSE}
# not run in knitted doc
rm(list = ls())
.rs.restartR()
```

```{r "load libraries", echo = FALSE, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(plotly)
library(knitr)
library(readxl)
library(tidyr)
source("functions.R")
opts_chunk$set(dev=c("png", "pdf"), dev.args=list(pdf = list(encoding="WinAnsi", useDingbats=FALSE)),
  fig.keep="all", fig.path = file.path("plots", "smoky_hollows_july")) ### CHANGE PLOT PREFIX ###
```


## Load data - LOW CONCENTRATION

```{r}
raw <- 
  bind_rows(
    read_excel(file.path("data", "20170517low_ND_data_export.xlsx")) 
  ) %>% 
  select(File, Nr., Start, Rt, End, # filename and retention times
         Ampl44=`Ampl 44`, Ampl45=`Ampl 45`, # amplitudes
         Area44=`rIntensity 44`, Area45=`rIntensity 45`, # areas
         Intensity=`Intensity All`, #peak intensities
         R13C = `R 13C/12C`, d13C = `13C/12C` # ratio and delta values
  )
```

```{r, echo=FALSE, eval=FALSE}
# file names
raw$File %>% unique() %>%  paste(collapse = "\n") %>% message()
```


### Map peaks

```{r}
### CHANGE MAPPING FILE NAME ###
data_matched <- map_peaks(raw, metadata_file = file.path("metadata", "old_standard_peaks_July20.xlsx"), quiet = FALSE)
```

### Add standard values

```{r}
standards <- read_excel(file.path("metadata", "gc_irms_indiana_A6.xlsx"))
data_w_stds <- data_matched %>% 
  filter(type == "standard", is_ref_peak == "no") %>% 
  left_join(standards, by = "compound") %>% 
  mutate(is_std = !is.na(true.d13C) | !is.na(true.d2H))
```

## Evaluate standards

### Visualize

```{r, fig.width = 9, fig.height=6}
data_w_stds %>% 
  ggplot() +
  aes(x = true.d13C, y = d13C, color = File) + 
  geom_smooth(method = "lm", se = FALSE, alpha = 0.5) +
  geom_point() +
  theme_bw() +
  theme(legend.position = "none") 
```

### Overview

```{r}
std_corrections <- 
  data_w_stds %>% 
  group_by(File) %>% 
  do({
    m <- lm(d13C ~ true.d13C, data = .)
    data_frame(
      offset = coefficients(m)[1],
      slope = coefficients(m)[2],
      delta_ref_CO = -offset/slope,
      max_residual = max(summary(m)$residuals),
      rss = sum(summary(m)$residuals^2),
      r2 = summary(m)$r.squared
    )
  }) 
std_corrections %>% knitr::kable(d = 3)
```


# isotope values check by rentention time
```{r, fig.width = 9, fig.height=6}
data_w_stds %>% 
  ggplot() +
  aes(x = Rt, y = d13C, color = File) + 
  geom_point() +
  theme_bw() +
  theme(legend.position = "none") 
```

```{r}
ggplotly(ggplot2::last_plot() + theme(legend.position = "none"))
```


# peak intensities check by rentention time
```{r, fig.width = 9, fig.height=6}
data_w_stds %>% 
  ggplot() +
  aes(x = Rt, y = Intensity, color = File) + 
  geom_point() +
  theme_bw() +
  theme(legend.position = "none") 
```

```{r}
ggplotly(ggplot2::last_plot() + theme(legend.position = "none"))
```


## Load data - HIGH CONCENTRATION

```{r}
raw <- 
  bind_rows(
    read_excel(file.path("data", "20170519high_ND_data_export.xlsx")) 
  ) %>% 
  select(File, Nr., Start, Rt, End, # filename and retention times
         Ampl44=`Ampl 44`, Ampl45=`Ampl 45`, # amplitudes
         Area44=`rIntensity 44`, Area45=`rIntensity 45`, # areas
         Intensity=`Intensity All`, #peak intensities
         R13C = `R 13C/12C`, d13C = `13C/12C` # ratio and delta values
  )
```

```{r, echo=FALSE, eval=FALSE}
# file names
raw$File %>% unique() %>%  paste(collapse = "\n") %>% message()
```


### Map peaks

```{r}
### CHANGE MAPPING FILE NAME ###
data_matched <- map_peaks(raw, metadata_file = file.path("metadata", "old_standard_peaks_high_July20.xlsx"), quiet = FALSE)
```

### Add standard values

```{r}
standards <- read_excel(file.path("metadata", "gc_irms_indiana_A6.xlsx"))
data_w_stds <- data_matched %>% 
  filter(type == "standard", is_ref_peak == "no") %>% 
  left_join(standards, by = "compound") %>% 
  mutate(is_std = !is.na(true.d13C) | !is.na(true.d2H))
```

## Evaluate standards

### Visualize

```{r, fig.width = 9, fig.height=6}
data_w_stds %>% 
  ggplot() +
  aes(x = true.d13C, y = d13C, color = File) + 
  geom_smooth(method = "lm", se = FALSE, alpha = 0.5) +
  geom_point() +
  theme_bw() +
  theme(legend.position = "none") 
```

### Overview

```{r}
std_corrections <- 
  data_w_stds %>% 
  group_by(File) %>% 
  do({
    m <- lm(d13C ~ true.d13C, data = .)
    data_frame(
      offset = coefficients(m)[1],
      slope = coefficients(m)[2],
      delta_ref_CO = -offset/slope,
      max_residual = max(summary(m)$residuals),
      rss = sum(summary(m)$residuals^2),
      r2 = summary(m)$r.squared
    )
  }) 
std_corrections %>% knitr::kable(d = 3)
```


# isotope values check by rentention time
```{r, fig.width = 9, fig.height=6}
data_w_stds %>% 
  ggplot() +
  aes(x = Rt, y = d13C, color = File) + 
  geom_point() +
  theme_bw() +
  theme(legend.position = "none") 
```




# peak intensities check by rentention time
```{r, fig.width = 9, fig.height=6}
data_w_stds %>% 
  ggplot() +
  aes(x = Rt, y = Intensity, color = File) + 
  geom_point() +
  theme_bw() +
  theme(legend.position = "none") 
```

```{r}
ggplotly(ggplot2::last_plot() + theme(legend.position = "none"))
```


## Load data - VERY HIGH CONCENTRATION #1

```{r}
raw <- 
  bind_rows(
    read_excel(file.path("data", "20170526veryhigh_ND_data_export.xlsx")) 
  ) %>% 
  select(File, Nr., Start, Rt, End, # filename and retention times
         Ampl44=`Ampl 44`, Ampl45=`Ampl 45`, # amplitudes
         Area44=`rIntensity 44`, Area45=`rIntensity 45`, # areas
         Intensity=`Intensity All`, #peak intensities
         R13C = `R 13C/12C`, d13C = `13C/12C` # ratio and delta values
  )
```

```{r, echo=FALSE, eval=FALSE}
# file names
raw$File %>% unique() %>%  paste(collapse = "\n") %>% message()
```

### Map peaks

```{r}
### CHANGE MAPPING FILE NAME ###
data_matched <- map_peaks(raw, metadata_file = file.path("metadata", "old_standard_peaks_veryhigh_July20.xlsx"), quiet = FALSE)
```

### Add standard values

```{r}
standards <- read_excel(file.path("metadata", "gc_irms_indiana_A6.xlsx"))
data_w_stds <- data_matched %>% 
  filter(type == "standard", is_ref_peak == "no") %>% 
  left_join(standards, by = "compound") %>% 
  mutate(is_std = !is.na(true.d13C) | !is.na(true.d2H))
```

## Evaluate standards

### Visualize

```{r, fig.width = 9, fig.height=6}
data_w_stds %>% 
  ggplot() +
  aes(x = true.d13C, y = d13C, color = File) + 
  geom_smooth(method = "lm", se = FALSE, alpha = 0.5) +
  geom_point() +
  theme_bw() +
  theme(legend.position = "none") 
```

### Overview

```{r}
std_corrections <- 
  data_w_stds %>% 
  group_by(File) %>% 
  do({
    m <- lm(d13C ~ true.d13C, data = .)
    data_frame(
      offset = coefficients(m)[1],
      slope = coefficients(m)[2],
      delta_ref_CO = -offset/slope,
      max_residual = max(summary(m)$residuals),
      rss = sum(summary(m)$residuals^2),
      r2 = summary(m)$r.squared
    )
  }) 
std_corrections %>% knitr::kable(d = 3)
```


# isotope values check by rentention time
```{r, fig.width = 9, fig.height=6}
data_w_stds %>% 
  ggplot() +
  aes(x = Rt, y = d13C, color = File) + 
  geom_point() +
  theme_bw() +
  theme(legend.position = "none") 
```

```{r}
ggplotly(ggplot2::last_plot() + theme(legend.position = "none"))
```


# peak intensities check by rentention time
```{r, fig.width = 9, fig.height=6}
data_w_stds %>% 
  ggplot() +
  aes(x = Rt, y = Intensity, color = File) + 
  geom_point() +
  theme_bw() +
  theme(legend.position = "none") 
```

```{r}
ggplotly(ggplot2::last_plot() + theme(legend.position = "none"))
```


## Load data - VERY HIGH CONCENTRATION #2

```{r}
raw <- 
  bind_rows(
    read_excel(file.path("data", "20170601veryhigh_ND_data_export.xlsx")) 
  ) %>% 
  select(File, Nr., Start, Rt, End, # filename and retention times
         Ampl44=`Ampl 44`, Ampl45=`Ampl 45`, # amplitudes
         Area44=`rIntensity 44`, Area45=`rIntensity 45`, # areas
         Intensity=`Intensity All`, #peak intensities
         R13C = `R 13C/12C`, d13C = `13C/12C` # ratio and delta values
  )
```

```{r, echo=FALSE, eval=FALSE}
# file names
raw$File %>% unique() %>%  paste(collapse = "\n") %>% message()
```


### Map peaks

```{r}
### CHANGE MAPPING FILE NAME ###
data_matched <- map_peaks(raw, metadata_file = file.path("metadata", "old_standard_peaks_veryhigh2_July20.xlsx"), quiet = FALSE)
```

### Add standard values

```{r}
standards <- read_excel(file.path("metadata", "gc_irms_indiana_A6.xlsx"))
data_w_stds <- data_matched %>% 
  filter(type == "standard", is_ref_peak == "no") %>% 
  left_join(standards, by = "compound") %>% 
  mutate(is_std = !is.na(true.d13C) | !is.na(true.d2H))
```

## Evaluate standards

### Visualize

```{r, fig.width = 9, fig.height=6}
data_w_stds %>% 
  ggplot() +
  aes(x = true.d13C, y = d13C, color = File) + 
  geom_smooth(method = "lm", se = FALSE, alpha = 0.5) +
  geom_point() +
  theme_bw() +
  theme(legend.position = "none") 
```

### Overview

```{r}
std_corrections <- 
  data_w_stds %>% 
  group_by(File) %>% 
  do({
    m <- lm(d13C ~ true.d13C, data = .)
    data_frame(
      offset = coefficients(m)[1],
      slope = coefficients(m)[2],
      delta_ref_CO = -offset/slope,
      max_residual = max(summary(m)$residuals),
      rss = sum(summary(m)$residuals^2),
      r2 = summary(m)$r.squared
    )
  }) 
std_corrections %>% knitr::kable(d = 3)
```


# isotope values check by rentention time
```{r, fig.width = 9, fig.height=6}
data_w_stds %>% 
  ggplot() +
  aes(x = Rt, y = d13C, color = File) + 
  geom_point() +
  theme_bw() +
  theme(legend.position = "none") 
```

```{r}
ggplotly(ggplot2::last_plot() + theme(legend.position = "none"))
```


# peak intensities check by rentention time
```{r, fig.width = 9, fig.height=6}
data_w_stds %>% 
  ggplot() +
  aes(x = Rt, y = Intensity, color = File) + 
  geom_point() +
  theme_bw() +
  theme(legend.position = "none") 
```

```{r}
ggplotly(ggplot2::last_plot() + theme(legend.position = "none"))
```

